---
title: "Code for the paper by Dean NE, Halloran B, and Zarnitsyna VI, Vaccine 2024"
author: "Veronika Zarnitsyna"
date: "11/10/2024"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


**You can click the 'Knit' button to generate a PDF file with all the figures or run the corresponding chunks for each figure separately.** 

**Figure 1 Panels C and D.** The vaccine effectiveness of vaccines against Covid-19. The data were digitized using WebPlotDigitizer (https://automeris.io/) from Lin et al., NEJM 2022 (PMID: 35020982) and added as two first columns of data in files "mRNA.csv" and "BNT162b2.csv". 

```{r figure1CD, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}

rm(list=ls()) 
library(readr)   # Read tabular data.
library(tidyr)   # Data frame tidying functions.
library(dplyr)   # General data frame manipulation.
library(ggplot2) # Flexible plotting.
library(cowplot) # to have multiple ggplots
#library(griddExtra) # to have multiple panels with ggplot
library(data.table) # for as.data.table() function
require(scales) #to use 10^6 instead of 1e_06 on y-axis

#reading the data 
mRNA <- read.csv("mRNA.csv", header=TRUE)
BNT162b2 <- read.csv("BNT162b2.csv", header=TRUE)

#The third column data in the data files was calculated as VEp=1-(1-mRNA$Hospitalization)/(1-mRNA$Covid) or VEp=1-(1-BNT162b2$Hospitalization)/(1-BNT162b2$Covid)

mRNA$Hospitalization=100*mRNA$Hospitalization
mRNA$Covid=100*mRNA$Covid
mRNA$VEp=100*mRNA$VEp
BNT162b2$Hospitalization=100*BNT162b2$Hospitalization
BNT162b2$Covid=100*BNT162b2$Covid
BNT162b2$VEp=100*BNT162b2$VEp

############################################################################################
My_Theme = theme(
  axis.title.x = element_text(size = 16),
  axis.text.x = element_text(size = 14),
  axis.title.y = element_text(size = 16),
  axis.text.y = element_text(size = 14))

print ("Figure 1 Panel C: Estimated VE components for mRNA-1273 vaccine")

mRNAplot <- ggplot(data =mRNA, aes(x = Time)) +
  geom_line(data = mRNA, aes(y = Hospitalization), linetype = "dashed", color = "red",size=1.) +      geom_line(data = mRNA, aes(y = Covid), linetype = "dotted",color = "red", size=1.) +
  geom_line(data = mRNA, aes(y = VEp), linetype = "solid", color = "red", size=1.1) + My_Theme

# Print the plot to include it in the document
print(mRNAplot)
ggsave("mRNA.eps", plot=mRNAplot, width = 6, height = 4, device="eps")

print ("Figure 1 Panel D: Estimated VE components for BNT162b2 vaccine")

BNTplot <- ggplot(data =BNT162b2, aes(x = Time)) +
  geom_line(data = BNT162b2, aes(y = Hospitalization),color = "blue",linetype = "dashed", size=1.) +   geom_line(data = BNT162b2, aes(y = Covid), linetype = "dotted", color = "blue", size=1.) +
  geom_line(data = BNT162b2, aes(y = VEp), linetype = "solid", color = "blue", size=1.2) + My_Theme

# Print the plot to include it in the document
print(BNTplot)

#ggsave("BNT162b2.eps", plot=BNTplot, width = 6, height = 4, device="eps")
```

**Figure 3. Simulation results for default scenario with waning VEs.**

Note that the code generating the data is commented (above the line ### Figure 3 Panel A ###) as it may take several hours to generate the simulation results files from 100 simulations. If uncommented, it will generate 3 sets of 100 files, allowing for plotting the dynamics of susceptible individuals and bar plots for each simulation, similar to the closest-to-average simulation shown in Figure 3. It will also create a summary file for VEps from 100 simulations, 'VEpbasic_waning730f007_1.csv', which will be used for further analysis. The file 'VEpbasic_waning730f007_1.csv,' along with three files for the simulation closest to the average ('susceptible1_69.csv', 'barplotALL1_69.csv', 'barplotALLS1_69.csv'), are provided for further analysis and plotting the results, allowing you to skip running the actual simulations.

The generated file "VEpbasic_waning730f007_1.csv" will also be used as a baseline condition in the chunk for plotting Figure 5 below.

```{r figure3, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
rm(list=ls()) 
library(readr)   # Read tabular data.
library(tidyr)   # Data frame tidying functions.
library(dplyr)   # General data frame manipulation.
library(ggplot2) # Flexible plotting.

# VEmodelwithbarplots<-function(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m){ #m is the number of simulations
#   with(as.list(c(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m)),{ # m is the number of the simulation
# # population
# N =200000 #number of unvaccinated people      
# NV=200000 #number of vaccinated people 
# #Unvaccinated
# infection_force = 1/180.   #4./180. # 30% attack rate over 180 days; to be adjusted
# infection_to_serious_force = 1/10. # base chance to get severe disease if infected
# #Vaccinated
# infection_forceV = 1/180. #4./180
# infection_to_serious_forceV = 1/10. # base chance to get severe disease if infected while being vaccinated
# 
# if (VEswaning>0) {
# VEs_waining= (alpha2-alpha4)/VEswaning #(alpha2-alpha4)/180.
# }
# if (VEpwaning>0) {
# VEp_waining= (beta2-beta4)/VEpwaning 
# }
# 
# group1=ceiling(N*frac1)-1 #correction
# group3=ceiling(NV*frac2)-1 #correction
# 
# ###########################block for unvaccinated##################################
# #initializing groups S and SS with different protection
# S = 0.0*(1:N)
# SS = 0.0*(1:N)
# infect=0*(1:N)
# ser=0*(1:N) #severe cases
# 
# #two groups: 'vulnerable'(group1) and 'immunocompetent'
# S[1:group1]=alpha1 # protection against infection in unvaccinated: 1 means base, <1 means more protected
# S[(group1+1):N]=alpha2
# SS[1:group1]=beta1 # protection against severe disease: 1 means base, <1 means more protected
# SS[(group1+1):N]=beta2 
# 
# for (i in 1:180) #i corresponds to the day, j corresponds to the individal/person
# { # cycle per days
#   for (j in 1:N) # cycle per people
#   {
#     if ((infect[j]==0) && (infection_force*S[j] > runif(1))) # infection probability
#     {
#       infect[j]=i #day of infection
#       if (infection_to_serious_force*SS[j] > runif(1)) # progression to severe probability
#       {
#         ser[j]=i #day of becoming severely ill
#       }
#     }
#   }
# }
# # sum_infect=sum(infect>0)
# # sum_ser= sum(ser>0)
# # ratio_inf_to_ser = sum_infect/sum_ser
# 
# #calculate at what month each person was infected; 0 - person is not infected
# infect_month=(infect+30)%/%31 
# 
# a=data.frame(S,infect,infect_month,ser)
# b = a %>% group_by(infect) %>% summarise(mean=mean(S))
# 
# #infect_month=0 in 'c' shows the line for uninfected; if absent -all infected; if '6' is absent - noone infected during month 6
# c = a %>% group_by(infect_month) %>% summarise(mean_per_Month=mean(S),N=n())
# 
# ser_month=(ser+30)%/%30.1
# a_ss=data.frame(S,ser,ser_month)
# ##ser_month=0 in 'd' shows the line for those not progressed to severe disease 
# d = a_ss %>% group_by(ser_month) %>% summarise(SS_per_Month=mean(S),Nser=n())
# 
# ###############################################vaccinated##############################
# #initializing groups S and SS with different protection
# SV = 0.0*(1:NV)
# SSV = 0.0*(1:NV)
# infectV=0*(1:NV)
# serV=0*(1:NV)
# 
# #two groups: 'vulnerable'(group3) and 'immunocompetent'in vaccinated
# SV[1:group3]=alpha3 # protection against infection in vaccinated: 1 means base, <1 means more protected
# SSV[1:group3]=beta3 # protection against severe disease in vaccinated: 1 means base, <1 means more protected
# 
# if (VEswaning==0) {
# # #two lines below are needed only for stable progression to severe disease#
# # #also need to comment the line SSV=SV below
#   SV[(group3+1):NV]=alpha4
# }
# 
# if (VEpwaning==0) {
# # #two lines below are needed only for stable progression to severe disease#
# # #also need to comment the line SSV=SV below
#   SSV[(group3+1):NV]=beta4
# }
# 
# for (i in 1:180)  { # cycle per days
#   if (VEswaning>0) {
#   SV[(group3+1):NV]=alpha4
#   alpha4=min(alpha2,alpha4+VEs_waining)
#   }
#   if (VEpwaning>0) {
#   SSV[(group3+1):NV]=beta4
#   beta4=min(beta2,beta4+VEp_waining)
#   }
# 
#   for (j in 1:N) # cycle per people
#   {
#     if ((infectV[j]==0) && (infection_forceV*SV[j] > runif(1))) # infection probability
#     {
#       infectV[j]=i
#       if (infection_to_serious_forceV*SSV[j] > runif(1)) # serious probability
#       {
#         serV[j]=i
#       }
#     }
#   }
# }
# sum_infectV=sum(infectV>0)
# sum_serV= sum(serV>0)
# ratio_inf_to_serV = sum(infectV>0)/sum(serV>0)
# 
# infect_monthV=(infectV+30)%/%31
# aV=data.frame(SV,infectV,infect_monthV,serV)
# bV = aV %>% group_by(infectV) %>% summarise(mean=mean(SV))
# cV = aV %>% group_by(infect_monthV) %>% summarise(mean_per_MonthV=mean(SV),NV=n())
#  
# ser_monthV=(serV+30)%/%30.1
# a_ssV=data.frame(SV,serV,ser_monthV)
# dV = a_ssV %>% group_by(ser_monthV) %>% summarise(SS_per_MonthV=mean(SV),NserV=n())
# 
# #########generate file to report susceptible left######
# susceptible=data.frame(S,infect,infect_month,ser,SV,infectV,infect_monthV,serV)
# write.csv(susceptible, file = paste0("susceptible1_",m,".csv"), row.names = FALSE)
# ###############################################################################################
# ######################################BAR PLOTS################################################
# # adding labels
# SG = rep('', times=N) # group label
# SG[1:group1]='1' # 
# SG[(group1+1):N]='2'
# 
# a_per_group=data.frame(SG,S,infect_month)
# c_per_group = a_per_group %>%
#   filter(infect_month > 0 ) %>%
#   group_by(SG,infect_month) %>% 
#   summarise(mean_per_Month=mean(S),N_inf=n())
# # plotting
# barplotU = data.frame(G = c_per_group$SG ,Month = c_per_group$infect_month,N_inf = c_per_group$N_inf)
# 
# SVG = rep('', times=N) #group label
# SVG[1:group3]="3"
# SVG[(group3+1):NV]="4"
# 
# ##############################Bar Plot for Infected##################################
# av_per_group=data.frame(SVG,SV,infect_monthV)
# cv_per_group = av_per_group %>%
#   filter(infect_monthV > 0 ) %>%
#   group_by(SVG,infect_monthV) %>% 
#   summarise(mean_per_Month=mean(SV),N_inf=n())
# barplotV = data.frame(G = cv_per_group$SVG ,Month = cv_per_group$infect_monthV,N_inf= cv_per_group$N_inf)
# 
# # all four groups combined for the bar plot
# barplotALL=rbind(barplotU,barplotV)
# #ggplot(barplotALL, aes(x = Month, y = N_inf, fill = G, colour = G)) + geom_bar(stat = "identity")
# #ggsave("barplotALL.eps", width = 6, height = 4.5, device="eps") 
# write.csv(barplotALL, file = paste0("barplotALL1_",m,".csv"), row.names = FALSE)
# 
# ##############################Bar Plot for Severe##################################
# #################progressed to severe in unvaccinated###################
# as_per_group=data.frame(SG,S,ser_month)
# cs_per_group = as_per_group %>%
#   filter(ser_month > 0 ) %>%
#   group_by(SG,ser_month) %>% 
#   summarise(mean_per_Month=mean(S),N_ser=n())
# barplotSU = data.frame(G = cs_per_group$SG ,Month = cs_per_group$ser_month,N_ser= cs_per_group$N_ser)
# #####progressed to severe in vaccinated###################
# avs_per_group=data.frame(SVG,SV,ser_monthV)
# cvs_per_group = avs_per_group %>%
#   filter(ser_monthV > 0 ) %>%
#   group_by(SVG,ser_monthV) %>% 
#   summarise(mean_per_Month=mean(SV),N_ser=n())
# barplotSV = data.frame(G = cvs_per_group$SVG ,Month = cvs_per_group$ser_monthV,N_ser= cvs_per_group$N_ser)
# 
# barplotALLS=rbind(barplotSU,barplotSV)
# write.csv(barplotALLS, file = paste0("barplotALLS1_",m,".csv"), row.names = FALSE)
# ###############################################################################################
# ##################################calculating VEp###############################
# VEp=1-(dV$NserV/cV$NV)[2:7]/(d$Nser/c$N)[2:7] 
#    return(VEp)
#    })
# }
# 
# set.seed(123) #setting seed keeps ALL 100 simulations the same if you repeat it
# simul_number=100
# VEpbasic <- matrix(NA,nrow=simul_number,ncol=6)
# for (m in 1:simul_number) {
# VEpbasic[m,] <- VEmodelwithbarplots(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,730,0,m)
# }
# write.csv(VEpbasic, file = "VEpbasic_waning730f007_1.csv", row.names = FALSE)

###########################  Figure 3 Panel A   ############################################
########Plotting all 100 simulation and the one that is closest to average####################

Plotting100<-function(filename){ 
  with(as.list(c(filename)),{
    
# Read the CSV file
file = paste0(filename, ".csv")
data <- read.csv(file)
  
strings <- data
# Calculate the overall average values 
overall_averaged_values <- apply(strings, 2, mean) 

# Function to find the string with the closest values to the overall averaged values 
find_closest_string <- function(strings, overall_averaged_values) { 
  # Calculate the distances between each string and the overall averaged values 
  distances <- apply(strings, 1, function(x) sqrt(sum((x - overall_averaged_values)^2))) 
  # Find the index of the string with the minimum distance 
  closest_string_index <- which.min(distances) 
  return(list(string = strings[closest_string_index, ], index = closest_string_index)) 
  } 
# Find the closest string to the overall averaged values 
closest_string <- find_closest_string(strings, overall_averaged_values) 

pdf(file = paste0(filename, ".pdf"), width = 6.7, height = 5.5, family = "Helvetica", pointsize = 12, onefile = TRUE)
# Plot all strings 
matplot(1:6, t(strings), type = "l", col = "grey", lty = 1, xlab = "Study Month", ylab = "VEp") 
# Plot the chosen closest string as a thick line 
lines(1:6, closest_string$string, type = "l", lwd = 2) 
#legend("topleft", legend = paste("Closest to Average (VEp#", closest_string$index,")", sep = " "), col = "black", lwd = 2)
dev.off()
     return(closest_string$index)
   })
 }

################finding the closest to averaged VE values for each condition ############ 
allfilenames <- c("VEpbasic_waning730f007_1")

for (k in 1:length(allfilenames)) {
index=Plotting100(allfilenames[k])
file = paste0(allfilenames[k], ".csv")
data=read.csv(file)

# Extract the specific row and convert it to a numeric vector
# Create the variable with the name in the name vector
#the result is a varibale with the name allfilenames[i] that has the simulation values closest to the 100 simulation average
assign(allfilenames[k], as.numeric(data[index,]), envir = .GlobalEnv)
}

print ("Figure 3 Panel A")
knitr::include_graphics("VEpbasic_waning730f007_1.pdf")

##################################### PANEL B ######################################
###############Plotting fractions of susceptibles for each grop##########

# Load the CSV file
data <- read.csv("susceptible1_69.csv")  

# Extract relevant columns (assuming 2nd and 6th columns contain infection dates)
infection_dates <- data[[2]]
infection_dates_vaccinated <- data[[6]]

# Define groups
group1 <- infection_dates[1:14001]
group2 <- infection_dates[14002:200000]
group3 <- infection_dates_vaccinated[1:14001]
group4 <- infection_dates_vaccinated[14002:200000]
total <- rbind(group1, group2, group3, group4)

# Function to calculate the fraction of susceptible people per day
fraction_susceptible <- function(group, days = 1:180) {
  total <- length(group)  # Total individuals in the group

  # Calculate the fraction of susceptible individuals per day
  sapply(days, function(day) {
    still_susceptible <- sum(group == 0 | group > day)  # Count individuals not infected by this day
    still_susceptible / total  # Fraction of susceptible individuals
  })
}

# Apply the function to each group
frac1 <- fraction_susceptible(group1)
frac2 <- fraction_susceptible(group2)
frac3 <- fraction_susceptible(group3)
frac4 <- fraction_susceptible(group4)

# Prepare data for plotting (include day 0 as 1 for all groups)
plot_data <- data.frame(
  Days = rep(0:180, 4),  # Days for x-axis positioning
  Fraction = c(1, frac1, 1, frac2, 1, frac3, 1, frac4),  # Fractions for each group
  Group = factor(rep(1:4, each = 181))  # Group labels
)

# Define color-blind-friendly colors for the groups
ColorBlindFriendly <- c("#E69F00", "#56B4E9", "#800080", "#009E73")

# Create the plot using ggplot2
ggplot(plot_data, aes(x = Days, y = Fraction, color = Group)) +
  geom_line(size = 1.5) +  # Make lines thicker with size parameter
  scale_colour_manual(values = ColorBlindFriendly) +
  scale_x_continuous(
    breaks = seq(0, 180, by = 30),  # Set breaks at every 30 days
    labels = c("0", "1", "2", "3", "4", "5", "6")  # Correct labels
  ) +
  labs(
    #title = "Fraction of Susceptible Individuals in Different Groups",
    x = "Study Month",  # X-axis label
    y = "Fraction of Susceptible",  # Y-axis label
    color = "Group"  # Legend title
  ) +
  theme(
    axis.title.x = element_text(size = 18),  # X-axis label size
    axis.title.y = element_text(size = 18),  # Y-axis label size
    axis.text.x = element_text(size = 18),   # X-axis tick label size
    axis.text.y = element_text(size = 18),   # Y-axis tick label size
    legend.title = element_text(size = 18),  # Legend title size
    legend.text = element_text(size = 18)    # Legend text size
  )
# Print the plot to include it in the document
print ("Figure 3 Panel B")
ggsave("Figure3PanelB.eps", width = 6, height = 4.5, device = "eps")

####################################### PanelC ########################################

barplotALL <- read.csv(file="barplotALL1_69.csv", row.names = NULL)

# Ensure 'Group' is a factor
barplotALL$Group <- as.factor(barplotALL$Group)

# Define colors for each group 
ColorBlindFriendly <- c("#E69F00", "#56B4E9", "#800080", "#009E73")
# Define darker versions of the colors for the borders
DarkerBorders <- c("#B36200", "#377CB3", "#4B0082", "#006F4D")

# Basic bar plot with color-blind-friendly palette and darker borders
ggplot(barplotALL, aes(x = factor(Month), y = N_inf, fill = Group, colour = Group)) +
  geom_bar(stat = "identity", width = 0.7, size = 0.8) +  # Bars with borders
  scale_fill_manual(values = ColorBlindFriendly) +  # Use color-blind-friendly fill colors
  scale_colour_manual(values = DarkerBorders) +  # Use darker colors for the borders
  theme(
    axis.text = element_text(size = 14),  # Increase size of axis tick numbers
    axis.title = element_text(size = 16),  # Increase size of axis titles
    legend.text = element_text(size = 12),  # Increase legend text size
    legend.title = element_text(size = 14)  # Increase legend title size
  ) + 
  labs(x = "Study Month", y = "Count of Infected")   # Adjust axis labels
print ("Figure 3 Panel C")
ggsave("Figure3PanelC.eps", width = 6, height = 4.5, device="eps")

####################################### Panel D ########################################
barplotALLS <- read.csv(file="barplotALLS1_69.csv", row.names = NULL)
barplotALLS$Group <- as.factor(barplotALLS$Group)

# Basic bar plot with grayscale colors (no alpha adjustment)
ggplot(barplotALLS, aes(x = factor(Month), y = N_ser, fill = Group, colour = Group)) +
 geom_bar(stat = "identity", width = 0.7, size = 0.8) +  # Bars with borders
  scale_fill_manual(values = ColorBlindFriendly) +  # Use color-blind-friendly fill colors
  scale_colour_manual(values = DarkerBorders) +  # Use darker colors for the borders
  theme(
    axis.text = element_text(size = 14),  # Increase size of axis tick numbers
    axis.title = element_text(size = 16),  # Increase size of axis titles
    legend.text = element_text(size = 12),  # Increase legend text size
    legend.title = element_text(size = 14)  # Increase legend title size
  ) + 
  labs(x = "Study Month", y = "Count of Severe")  # Adjust axis labels
print ("Figure 3 Panel D")
ggsave("Figure3PanelD.eps", width = 6, height = 4.5, device="eps")

################################################################################################
################################Calculating the attack rate#################################
################################################################################################
data1_69 <- read.csv("susceptible1_69.csv")  

# # Calculate the total number of 0s in columns 2 and 6

total_zeros1_69 <- sum(data1_69[[2]] == 0, na.rm = TRUE) + sum(data1_69[[6]] == 0, na.rm = TRUE)

#attack rates
AR1_69=(400000-total_zeros1_69)/400000 #0.4
print(AR1_69) #0.4
```

\newpage
**Figure 4. Simulation results for default scenario with constant VEs.**

Note that the code generating the data is commented (above the line ### Figure 4 Panel A ###) similar to the chunk for Figure 3 as it may take several hours to generate the simulation results files from 100 simulations. If uncommented, the structure of generated data will be similar to described above for Figure 3. Similarly to Figure 3, the file 'VEpbasic_waning730f007VEsconstant.csv' along with three files for the simulation closest to the average ('susceptibleVEsconstant38.csv', 'barplotALLVEsconstant38.csv', 'barplotALLSVEsconstant38.csv'), are provided for further analysis and plotting the results, allowing you to skip running the actual simulations.

```{r figure4, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}

rm(list=ls()) 
library(readr)   # Read tabular data.
library(tidyr)   # Data frame tidying functions.
library(dplyr)   # General data frame manipulation.
library(ggplot2) # Flexible plotting.

# VEmodelwithbarplots<-function(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m){ #m is the number of simulations
#   with(as.list(c(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m)),{ # m is the number of the simulation
# # population
# N =200000 #number of unvaccinated people      
# NV=200000 #number of vaccinated people 
# #Unvaccinated
# infection_force = 1./180.   #4./180. # 30% attack rate over 180 days; to be adjusted
# infection_to_serious_force = 1/10. # base chance to get severe disease if infected
# #Vaccinated
# infection_forceV = 1./180. #4./180
# infection_to_serious_forceV = 1/10. # base chance to get severe disease if infected while being vaccinated
# 
# if (VEswaning>0) {
# VEs_waining= (alpha2-alpha4)/VEswaning #(alpha2-alpha4)/180.
# }
# if (VEpwaning>0) {
# VEp_waining= (beta2-beta4)/VEpwaning 
# }
# 
# group1=ceiling(N*frac1)
# group3=ceiling(NV*frac2)
# 
# ###########################block for unvaccinated##################################
# #initializing groups S and SS with different protection
# S = 0.0*(1:N)
# SS = 0.0*(1:N)
# infect=0*(1:N)
# ser=0*(1:N) #severe cases
# 
# #two groups: 'vulnerable'(group1) and 'immunocompetent'
# S[1:group1]=alpha1 # protection against infection in unvaccinated: 1 means base, <1 means more protected
# S[(group1+1):N]=alpha2
# SS[1:group1]=beta1 # protection against severe disease: 1 means base, <1 means more protected
# SS[(group1+1):N]=beta2 
# 
# for (i in 1:180) #i corresponds to the day, j corresponds to the individal/person
# { # cycle per days
#   for (j in 1:N) # cycle per people
#   {
#     if ((infect[j]==0) && (infection_force*S[j] > runif(1))) # infection probability
#     {
#       infect[j]=i #day of infection
#       if (infection_to_serious_force*SS[j] > runif(1)) # progression to severe probability
#       {
#         ser[j]=i #day of becoming severely ill
#       }
#     }
#   }
# }
# 
# #calculate at what month each person was infected; 0 - person is not infected
# infect_month=(infect+30)%/%31 
# 
# a=data.frame(S,infect,infect_month,ser)
# b = a %>% group_by(infect) %>% summarise(mean=mean(S))
# 
# #infect_month=0 in 'c' shows the line for uninfected; if absent -all infected; if '6' is absent - noone infected during month 6
# c = a %>% group_by(infect_month) %>% summarise(mean_per_Month=mean(S),N=n())
# 
# ser_month=(ser+30)%/%30.1
# a_ss=data.frame(S,ser,ser_month)
# ##ser_month=0 in 'd' shows the line for those not progressed to severe disease 
# d = a_ss %>% group_by(ser_month) %>% summarise(SS_per_Month=mean(S),Nser=n())
# 
# ###############################################vaccinated##############################
# #initializing groups S and SS with different protection
# SV = 0.0*(1:NV)
# SSV = 0.0*(1:NV)
# infectV=0*(1:NV)
# serV=0*(1:NV)
# 
# #two groups: 'vulnerable'(group3) and 'immunocompetent'in vaccinated
# SV[1:group3]=alpha3 # protection against infection in vaccinated: 1 means base, <1 means more protected
# SSV[1:group3]=beta3 # protection against severe disease in vaccinated: 1 means base, <1 means more protected
# 
# if (VEswaning==0) {
# # #two lines below are needed only for stable progression to severe disease#
# # #also need to comment the line SSV=SV below
#   SV[(group3+1):NV]=alpha4
# }
# 
# if (VEpwaning==0) {
# # #two lines below are needed only for stable progression to severe disease#
# # #also need to comment the line SSV=SV below
#   SSV[(group3+1):NV]=beta4
# }
# 
# for (i in 1:180)  { # cycle per days
#   if (VEswaning>0) {
#   SV[(group3+1):NV]=alpha4
#   alpha4=min(alpha2,alpha4+VEs_waining)
#   }
#   if (VEpwaning>0) {
#   SSV[(group3+1):NV]=beta4
#   beta4=min(beta2,beta4+VEp_waining)
#   }
# 
#   for (j in 1:N) # cycle per people
#   {
#     if ((infectV[j]==0) && (infection_forceV*SV[j] > runif(1))) # infection probability
#     {
#       infectV[j]=i
#       if (infection_to_serious_forceV*SSV[j] > runif(1)) # serious probability
#       {
#         serV[j]=i
#       }
#     }
#   }
# }
# sum_infectV=sum(infectV>0)
# sum_serV= sum(serV>0)
# ratio_inf_to_serV = sum(infectV>0)/sum(serV>0)
# 
# infect_monthV=(infectV+30)%/%31
# aV=data.frame(SV,infectV,infect_monthV,serV)
# bV = aV %>% group_by(infectV) %>% summarise(mean=mean(SV))
# cV = aV %>% group_by(infect_monthV) %>% summarise(mean_per_MonthV=mean(SV),NV=n())
#  
# ser_monthV=(serV+30)%/%30.1
# a_ssV=data.frame(SV,serV,ser_monthV)
# dV = a_ssV %>% group_by(ser_monthV) %>% summarise(SS_per_MonthV=mean(SV),NserV=n())
# 
# #########generate file to report susceptible left######
# susceptible=data.frame(S,infect,infect_month,ser,SV,infectV,infect_monthV,serV)
# write.csv(susceptible, file = paste0("susceptibleVEsconstant",m,".csv"), row.names = FALSE)
# ###############################################################################################
# ######################################BAR PLOTS################################################
# # adding labels
# SG = rep('', times=N) # group label
# SG[1:group1]='1' # 
# SG[(group1+1):N]='2'
# 
# a_per_group=data.frame(SG,S,infect_month)
# c_per_group = a_per_group %>%
#   filter(infect_month > 0 ) %>%
#   group_by(SG,infect_month) %>% 
#   summarise(mean_per_Month=mean(S),N_inf=n())
# # plotting
# barplotU = data.frame(G = c_per_group$SG ,Month = c_per_group$infect_month,N_inf = c_per_group$N_inf)
# 
# SVG = rep('', times=N) #group label
# SVG[1:group3]="3"
# SVG[(group3+1):NV]="4"
# 
# ##############################Bar Plot for Infected##################################
# av_per_group=data.frame(SVG,SV,infect_monthV)
# cv_per_group = av_per_group %>%
#   filter(infect_monthV > 0 ) %>%
#   group_by(SVG,infect_monthV) %>% 
#   summarise(mean_per_Month=mean(SV),N_inf=n())
# barplotV = data.frame(G = cv_per_group$SVG ,Month = cv_per_group$infect_monthV,N_inf= cv_per_group$N_inf)
# 
# # all four groups combined for the bar plot
# barplotALL=rbind(barplotU,barplotV)
# write.csv(barplotALL, file = paste0("barplotALLVEsconstant",m,".csv"), row.names = FALSE)
# 
# ##############################Bar Plot for Severe##################################
# #################progressed to severe in unvaccinated###################
# as_per_group=data.frame(SG,S,ser_month)
# cs_per_group = as_per_group %>%
#   filter(ser_month > 0 ) %>%
#   group_by(SG,ser_month) %>% 
#   summarise(mean_per_Month=mean(S),N_ser=n())
# barplotSU = data.frame(G = cs_per_group$SG ,Month = cs_per_group$ser_month,N_ser= cs_per_group$N_ser)
# #####progressed to severe in vaccinated###################
# avs_per_group=data.frame(SVG,SV,ser_monthV)
# cvs_per_group = avs_per_group %>%
#   filter(ser_monthV > 0 ) %>%
#   group_by(SVG,ser_monthV) %>% 
#   summarise(mean_per_Month=mean(SV),N_ser=n())
# barplotSV = data.frame(G = cvs_per_group$SVG ,Month = cvs_per_group$ser_monthV,N_ser= cvs_per_group$N_ser)
# 
# barplotALLS=rbind(barplotSU,barplotSV)
# write.csv(barplotALLS, file = paste0("barplotALLSVEsconstant",m,".csv"), row.names = FALSE)
# ###############################################################################################
# ##################################calculating VEp###############################
# VEp=1-(dV$NserV/cV$NV)[2:7]/(d$Nser/c$N)[2:7] 
#    return(VEp)
#    })
# }
# 
# set.seed(123) #setting seed keeps ALL 100 simulations the same if you repeat it
# simul_number=100
# VEpbasic <- matrix(NA,nrow=simul_number,ncol=6)
# for (m in 1:simul_number) {
# VEpbasic[m,] <- VEmodelwithbarplots(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,0,0,m)
# }
# write.csv(VEpbasic, file = "VEpbasic_waning730f007VEsconstant.csv", row.names = FALSE)

#################  UNCOMMENT LINES ABOVE TO GENERATE THE DATA FILEs ########################

###########################  Figure 4 Panel A   ############################################
#Finding the closest to the average curve and it's number

########Plotting all 100 simulation and the one that is closest to average####################

Plotting100<-function(filename){ 
  with(as.list(c(filename)),{
    
# Read the CSV file
file = paste0(filename, ".csv")
data <- read.csv(file)
  
strings <- data
# Calculate the overall average values 
overall_averaged_values <- apply(strings, 2, mean) 

# Function to find the string with the closest values to the overall averaged values 
find_closest_string <- function(strings, overall_averaged_values) { 
  # Calculate the distances between each string and the overall averaged values 
  distances <- apply(strings, 1, function(x) sqrt(sum((x - overall_averaged_values)^2))) 
  # Find the index of the string with the minimum distance 
  closest_string_index <- which.min(distances) 
  return(list(string = strings[closest_string_index, ], index = closest_string_index)) 
  } 
# Find the closest string to the overall averaged values 
closest_string <- find_closest_string(strings, overall_averaged_values) 

pdf(file = paste0(filename, ".pdf"), width = 6.7, height = 5.5, family = "Helvetica", pointsize = 12, onefile = TRUE)
# Plot all strings 
matplot(1:6, t(strings), type = "l", col = "grey", lty = 1, xlab = "Study Month", ylab = "VEp") 
# Plot the chosen closest string as a thick line 
lines(1:6, closest_string$string, type = "l", lwd = 2) 
#legend("topleft", legend = paste("Closest to Average (VEp#", closest_string$index,")", sep = " "), col = "black", lwd = 2)
dev.off()
     return(closest_string$index)
   })
 }

################finding the closest to averaged VE values for each condition ############ 
allfilenames <- c("VEpbasic_waning730f007VEsconstant")

for (k in 1:length(allfilenames)) {
index=Plotting100(allfilenames[k])
file = paste0(allfilenames[k], ".csv")
data=read.csv(file)

# Extract the specific row and convert it to a numeric vector
# Create the variable with the name in the name vector
#the result is a varibale with the name allfilenames[i] that has the simulation values closest to the 100 simulation average
assign(allfilenames[k], as.numeric(data[index,]), envir = .GlobalEnv)
}

print ("Figure 4 Panel A")
knitr::include_graphics("VEpbasic_waning730f007VEsconstant.pdf")
###################################  Figure 4 Panel B   ################################
# Load the CSV file
data <- read.csv("susceptibleVEsconstant38.csv") 

# Extract relevant columns (assuming 2nd and 6th columns contain infection dates)
infection_dates <- data[[2]]
infection_dates_vaccinated <- data[[6]]

# Define groups
group1 <- infection_dates[1:14001]
group2 <- infection_dates[14002:200000]
group3 <- infection_dates_vaccinated[1:14001]
group4 <- infection_dates_vaccinated[14002:200000]
total <- rbind(group1, group2, group3, group4)

# Function to calculate the fraction of susceptible people per day
fraction_susceptible <- function(group, days = 1:180) {
  total <- length(group)  # Total individuals in the group

  # Calculate the fraction of susceptible individuals per day
  sapply(days, function(day) {
    still_susceptible <- sum(group == 0 | group > day)  # Count individuals not infected by this day
    still_susceptible / total  # Fraction of susceptible individuals
  })
}

# Apply the function to each group
frac1 <- fraction_susceptible(group1)
frac2 <- fraction_susceptible(group2)
frac3 <- fraction_susceptible(group3)
frac4 <- fraction_susceptible(group4)

# Prepare data for plotting (include day 0 as 1 for all groups)
plot_data <- data.frame(
  Days = rep(0:180, 4),  # Days for x-axis positioning
  Fraction = c(1, frac1, 1, frac2, 1, frac3, 1, frac4),  # Fractions for each group
  Group = factor(rep(1:4, each = 181))  # Group labels
)

# Define color-blind-friendly colors for the groups
ColorBlindFriendly <- c("#E69F00", "#56B4E9", "#800080", "#009E73")

# Create the plot using ggplot2
ggplot(plot_data, aes(x = Days, y = Fraction, color = Group)) +
  geom_line(size = 1.5) +  # Make lines thicker with size parameter
  scale_colour_manual(values = ColorBlindFriendly) +
  scale_x_continuous(
    breaks = seq(0, 180, by = 30),  # Set breaks at every 30 days
    labels = c("0", "1", "2", "3", "4", "5", "6")  # Correct labels
  ) +
  labs(
    #title = "Fraction of Susceptible Individuals in Different Groups",
    x = "Study Month",  # X-axis label
    y = "Fraction of Susceptible",  # Y-axis label
    color = "Group"  # Legend title
  ) +
  theme(
    axis.title.x = element_text(size = 18),  # X-axis label size
    axis.title.y = element_text(size = 18),  # Y-axis label size
    axis.text.x = element_text(size = 18),   # X-axis tick label size
    axis.text.y = element_text(size = 18),   # Y-axis tick label size
    legend.title = element_text(size = 18),  # Legend title size
    legend.text = element_text(size = 18)    # Legend text size
  )
print ("Figure 4 Panel B")
ggsave("Figure4PanelB.eps", width = 6, height = 4.5, device = "eps")

####################################### PanelC ########################################

barplotALL <- read.csv(file="barplotALLVEsconstant38.csv", row.names = NULL)

# Ensure 'Group' is a factor
barplotALL$Group <- as.factor(barplotALL$Group)

# Define colors for each group 
ColorBlindFriendly <- c("#E69F00", "#56B4E9", "#800080", "#009E73")
# Define darker versions of the colors for the borders
DarkerBorders <- c("#B36200", "#377CB3", "#4B0082", "#006F4D")

# Basic bar plot with color-blind-friendly palette and darker borders
ggplot(barplotALL, aes(x = factor(Month), y = N_inf, fill = Group, colour = Group)) +
  geom_bar(stat = "identity", width = 0.7, size = 0.8) +  # Bars with borders
  scale_fill_manual(values = ColorBlindFriendly) +  # Use color-blind-friendly fill colors
  scale_colour_manual(values = DarkerBorders) +  # Use darker colors for the borders
  theme(
    axis.text = element_text(size = 14),  # Increase size of axis tick numbers
    axis.title = element_text(size = 16),  # Increase size of axis titles
    legend.text = element_text(size = 12),  # Increase legend text size
    legend.title = element_text(size = 14)  # Increase legend title size
  ) + 
  labs(x = "Study Month", y = "Count of Infected")   # Adjust axis labels
print ("Figure 4 Panel C")
ggsave("Figure4PanelC.eps", width = 6, height = 4.5, device="eps")

####################################### Panel D ########################################
barplotALLS <- read.csv(file="barplotALLSVEsconstant38.csv", row.names = NULL)
barplotALLS$Group <- as.factor(barplotALLS$Group)

# Basic bar plot with grayscale colors (no alpha adjustment)
ggplot(barplotALLS, aes(x = factor(Month), y = N_ser, fill = Group, colour = Group)) +
 geom_bar(stat = "identity", width = 0.7, size = 0.8) +  # Bars with borders
  scale_fill_manual(values = ColorBlindFriendly) +  # Use color-blind-friendly fill colors
  scale_colour_manual(values = DarkerBorders) +  # Use darker colors for the borders
  theme(
    axis.text = element_text(size = 14),  # Increase size of axis tick numbers
    axis.title = element_text(size = 16),  # Increase size of axis titles
    legend.text = element_text(size = 12),  # Increase legend text size
    legend.title = element_text(size = 14)  # Increase legend title size
  ) + 
  labs(x = "Study Month", y = "Count of Severe")  # Adjust axis labels
print ("Figure 4 Panel D")
ggsave("Figure4PanelD.eps", width = 6, height = 4.5, device="eps")

```

**Figure 5. Parameter sensitivity analysis and its effect on observed VEp.**

This chunk will first generate .csv files with the following names: "VEpbasic_waning730f007_1.csv", "VEp4xfaster.csv", "VEp2xslower.csv", "VEp4xslower.csv", "VEpf001.csv", "VEpf015.csv", "VEpf03.csv", "VEpbeta2_1.csv", "VEpbeta2_01.csv", "VEpalpha4_001.csv", "VEpalpha4_025.csv", "VEpalpha4_05.csv", "VEpalpha4_1.csv", "VEpbeta4_001.csv", "VEpbeta4_025.csv", "VEpbeta4_05.csv", "VEpconstantVEs.csv", "VEpwaning180.csv", "VEpwaning365.csv", "VEpwaning730.csv". These data files will have 100 simulations for each indicated condition and will be processed to find the closest to average simulation that will be used for panels in Figure 5.

Note that generating these files may take a long time to complete (depending on the computer, it can take several days), so the part that generates these files is currently commented out. All .csv files that it would generate are provided for further analysis.

```{r figure5, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}

rm(list=ls()) 
library(readr)   # Read tabular data.
library(tidyr)   # Data frame tidying functions.
library(dplyr)   # General data frame manipulation.
library(ggplot2) # Flexible plotting.

# #########################  The part below is commented out #############################
# VEmodel<-function(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning){
#  with(as.list(c(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning)),{

# # population
# N =200000
# NV=200000

#########################FOR REDUCED ATTACK RATE######################
#Unvaccinated
# infection_force = 1./180.   #2./180. # 30% attack rate over 180 days; to be adjusted
# infection_to_serious_force = 1/10. # base chance to get severe disease while being vaccinated
# #Vaccinated
# infection_forceV = 1./180. # 30% base rate infection over 180 days to be adjusted
# infection_to_serious_forceV = 1/10. # base chance to get serious while being vaccinated
#
# if (VEswaning>0) {
# VEs_waining= (alpha2-alpha4)/VEswaning #(alpha2-alpha4)/180.
# }
# if (VEpwaning>0) {
# VEp_waining= (beta2-beta4)/VEpwaning #(alpha2-alpha4)/180.
# }
#
# group1=ceiling(N*frac1)
# group3=ceiling(NV*frac2)
#
# ####block for unvaccinated
# #initial groups
# S = 0.0*(1:N)
# SS = 0.0*(1:N)
# infect=0*(1:N)
# ser=0*(1:N) #severe cases
#
# S[1:group1]=alpha1 # susceptibility 1 - means base, while 0.64 below means more protected
# S[(group1+1):N]=alpha2
# SS[1:group1]=beta1 # serious susceptibility
# SS[(group1+1):N]=beta2 # serious susceptibility
#
# for (i in 1:180) #i corresponds to the day, j corresponds to the person
# { # cycle per days
#   for (j in 1:N) # cycle per people
#   {
#     if ((infect[j]==0) && (infection_force*S[j] > runif(1))) # infection probability
#     {
#       infect[j]=i #day of infection
#       if (infection_to_serious_force*SS[j] > runif(1)) # serious probability
#       {
#         ser[j]=i #day of becoming severely ill
#       }
#     }
#   }
# }
# sum_infect=sum(infect>0)
# sum_ser= sum(ser>0)
# ratio_inf_to_ser = sum(infect>0)/sum(ser>0)
#
# infect_month=(infect+30)%/%31
# a=data.frame(S,infect,infect_month,ser)
# b = a %>%
#   group_by(infect) %>%
#   summarise(mean=mean(S))
# c = a %>%
#   group_by(infect_month) %>%
#   summarise(mean_per_Month=mean(S),N=n())
#
# ser_month=(ser+30)%/%30.1
# a_ss=data.frame(S,ser,ser_month)
# d = a_ss %>%
#   group_by(ser_month) %>%
#   summarise(SS_per_Month=mean(S),Nser=n())
#
# ###############################################vaccinated##############################
# #initial groups
# SV = 0.0*(1:NV)
# SSV = 0.0*(1:NV)
# infectV=0*(1:NV)
# serV=0*(1:NV)
#
# SV[1:group3]=alpha3
# SSV[1:group3]=beta3
#
# if (VEswaning==0) {
# # #two lines below are needed only for stable progression to severe disease#
# # #also need to comment the line SSV=SV below
#   SV[(group3+1):NV]=alpha4
# }
#
# if (VEpwaning==0) {
# # #two lines below are needed only for stable progression to severe disease#
# # #also need to comment the line SSV=SV below
#   SSV[(group3+1):NV]=beta4
# }
#
# for (i in 1:180)  { # cycle per days
#   if (VEswaning>0) {
#   SV[(group3+1):NV]=alpha4
#   alpha4=min(alpha2,alpha4+VEs_waining)
#   }
#   if (VEpwaning>0) {
#   SSV[(group3+1):NV]=beta4
#   beta4=min(beta2,beta4+VEp_waining)
#   }
#
#   for (j in 1:N) # cycle per people
#   {
#     if ((infectV[j]==0) && (infection_forceV*SV[j] > runif(1))) # infection probability
#     {
#       infectV[j]=i
#       if (infection_to_serious_forceV*SSV[j] > runif(1)) # serious probability
#       {
#         serV[j]=i
#       }
#     }
#   }
# }
# sum_infectV=sum(infectV>0)
# sum_serV= sum(serV>0)
# ratio_inf_to_serV = sum(infectV>0)/sum(serV>0)
#
# infect_monthV=(infectV+30)%/%31
# aV=data.frame(SV,infectV,infect_monthV,serV)
# bV = aV %>%
#   group_by(infectV) %>%
#   summarise(mean=mean(SV))
# cV = aV %>%
#   group_by(infect_monthV) %>%
#   summarise(mean_per_MonthV=mean(SV),NV=n())
#
# ser_monthV=(serV+30)%/%30.1
# a_ssV=data.frame(SV,serV,ser_monthV)
# dV = a_ssV %>%
#   group_by(ser_monthV) %>%
#   summarise(SS_per_MonthV=mean(SV),NserV=n())
#
# #################calculating VEp###############################
# VEp=1-(dV$NserV/cV$NV)[2:7]/(d$Nser/c$N)[2:7]
#      return(VEp)
#    })
#  }
#
# ############Plotting 6 panels figure for Short Communication############
# # #initializing data matrix
# simul_number=100
# VEp4xfaster <- matrix(NA,nrow=simul_number,ncol=6)
# VEp2xslower <- matrix(NA,nrow=simul_number,ncol=6)
# VEp4xslower <- matrix(NA,nrow=simul_number,ncol=6)
# VEpf001 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpf015 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpf03 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpbeta2_1 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpbeta2_01 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpalpha4_001 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpalpha4_025 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpalpha4_05 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpalpha4_1 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpbeta4_001 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpbeta4_025 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpbeta4_05 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpconstantVEs <- matrix(NA,nrow=simul_number,ncol=6)
# VEpwaning180 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpwaning365 <- matrix(NA,nrow=simul_number,ncol=6)
# VEpwaning730 <- matrix(NA,nrow=simul_number,ncol=6)
#
# set.seed(123) #setting seed keeps ALL 100 simulations the same if you repeat it
# for (m in 1:simul_number) {
#
# VEp4xfaster[m,] <- VEmodel(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,182.5,0)
# VEp2xslower[m,] <- VEmodel(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,1460,0)
# VEp4xslower[m,] <- VEmodel(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,2920,0)
# print(m)
# VEpf001[m,] <- VEmodel(0.01,0.01,1,1,1,0.03,1,0.5,1,0.03,730,0)
# VEpf015[m,] <- VEmodel(0.15,0.15,1,1,1,0.03,1,0.5,1,0.03,730,0)
# VEpf03[m,]  <- VEmodel(0.3,0.3,1,1,1,0.03,1,0.5,1,0.03,730,0)
# print(m)
# VEpbeta2_1[m,] <- VEmodel(0.07,0.07,1,1,1,0.03,1,1,1,0.03,730,0)
# VEpbeta2_01[m,] <- VEmodel(0.07,0.07,1,1,1,0.03,1,0.1,1,0.03,730,0)
# VEpalpha4_001[m,] <- VEmodel(0.07,0.07,1,1,1,0.01,1,0.5,1,0.03,730,0)
# print(m)
# VEpalpha4_025[m,] <- VEmodel(0.07,0.07,1,1,1,0.25,1,0.5,1,0.03,730,0)
# VEpalpha4_05[m,] <- VEmodel(0.07,0.07,1,1,1,0.5,1,0.5,1,0.03,730,0)
# VEpalpha4_1[m,] <- VEmodel(0.07,0.07,1,1,1,1,1,0.5,1,0.03,730,0)
# print(m)
# VEpbeta4_001[m,] <- VEmodel(0.07,0.07,1,1,1,0.03,1,0.5,1,0.01,730,0)
# VEpbeta4_025[m,] <- VEmodel(0.07,0.07,1,1,1,0.03,1,0.5,1,0.25,730,0)
# VEpbeta4_05[m,]  <- VEmodel(0.07,0.07,1,1,1,0.03,1,0.5,1,0.5,730,0)
# VEpconstantVEs[m,] <- VEmodel(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,0,0)
# VEpwaning180[m,] <- VEmodel(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,730,180)
# VEpwaning365[m,] <- VEmodel(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,730,365)
# VEpwaning730[m,] <- VEmodel(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,730,730)
# print(m)
# }
# write.csv(VEp4xfaster, file = "VEp4xfaster.csv", row.names = FALSE)
# write.csv(VEp2xslower, file = "VEp2xslower.csv", row.names = FALSE)
# write.csv(VEp4xslower, file = "VEp4xslower.csv", row.names = FALSE)
# write.csv(VEpf001, file = "VEpf001.csv", row.names = FALSE)
# write.csv(VEpf015, file = "VEpf015.csv", row.names = FALSE)
# write.csv(VEpf03, file = "VEpf03.csv", row.names = FALSE)
# write.csv(VEpbeta2_1, file = "VEpbeta2_1.csv", row.names = FALSE)
# write.csv(VEpbeta2_01, file = "VEpbeta2_01.csv", row.names = FALSE)
# write.csv(VEpalpha4_001, file = "VEpalpha4_001.csv", row.names = FALSE)
# write.csv(VEpalpha4_025, file = "VEpalpha4_025.csv", row.names = FALSE)
# write.csv(VEpalpha4_05, file = "VEpalpha4_05.csv", row.names = FALSE)
# write.csv(VEpalpha4_1, file = "VEpalpha4_1.csv", row.names = FALSE)
# write.csv(VEpbeta4_001, file = "VEpbeta4_001.csv", row.names = FALSE)
# write.csv(VEpbeta4_025, file = "VEpbeta4_025.csv", row.names = FALSE)
# write.csv(VEpbeta4_05, file = "VEpbeta4_05.csv", row.names = FALSE)
# write.csv(VEpconstantVEs, file = "VEpconstantVEs.csv", row.names = FALSE)
# write.csv(VEpwaning180, file = "VEpwaning180.csv", row.names = FALSE)
# write.csv(VEpwaning365, file = "VEpwaning365.csv", row.names = FALSE)
# write.csv(VEpwaning730, file = "VEpwaning730.csv", row.names = FALSE)

#### UNCOMMENT THE LINES ABOVE IF YOU WOULD LIKE TO GENERATE ALL .csv FILES FROM SCRATCH #########

########Plotting all 100 simulation and the one that is closest to average##################
#it also create the cariable for VEp with the name of the file and data for closest to average

Plotting100<-function(filename){ 
  with(as.list(c(filename)),{
    
# Read the CSV file
file = paste0(filename, ".csv")
data <- read.csv(file)
  
strings <- data
# Calculate the overall average values 
overall_averaged_values <- apply(strings, 2, mean) 

# Function to find the string with the closest values to the overall averaged values 
find_closest_string <- function(strings, overall_averaged_values) { 
  # Calculate the distances between each string and the overall averaged values 
  distances <- apply(strings, 1, function(x) sqrt(sum((x - overall_averaged_values)^2))) 
  # Find the index of the string with the minimum distance 
  closest_string_index <- which.min(distances) 
  return(list(string = strings[closest_string_index, ], index = closest_string_index)) 
  } 
# Find the closest string to the overall averaged values 
closest_string <- find_closest_string(strings, overall_averaged_values) 

pdf(file = paste0(filename, ".pdf"), width = 6.7, height = 5.5, family = "Helvetica", pointsize = 12, onefile = TRUE)
# Plot all strings 
matplot(1:6, t(strings), type = "l", col = "grey", lty = 1, xlab = "Study Month", ylab = "VEp") 
# Plot the chosen closest string as a thick line 
lines(1:6, closest_string$string, type = "l", lwd = 2) 
#legend("topleft", legend = paste("Closest to Average (VEp#", closest_string$index,")", sep = " "), col = "black", lwd = 2)
dev.off()
     return(closest_string$index)
   })
 }

################finding the closest to averaged VE values for each condition ############ 
allfilenames <- c("VEpbasic_waning730f007_1","VEp4xfaster","VEp2xslower","VEp4xslower","VEpf001","VEpf015","VEpf03","VEpbeta2_1","VEpbeta2_01","VEpalpha4_001","VEpalpha4_025","VEpalpha4_05","VEpalpha4_1","VEpbeta4_001","VEpbeta4_025","VEpbeta4_05","VEpconstantVEs","VEpwaning180","VEpwaning365","VEpwaning730")

for (k in 1:length(allfilenames)) {
index=Plotting100(allfilenames[k])
file = paste0(allfilenames[k], ".csv")
data=read.csv(file)
# Extract the specific row and convert it to a numeric vector
# Create the variable with the name in the name vector
#the result is a varibale with the name allfilenames[i] that has the simulation values closest to the 100 simulation average
assign(allfilenames[k], as.numeric(data[index,]), envir = .GlobalEnv)
}
##########################################################################################

pdf(file = "Figure5.pdf", width = 12, height = 7, family = "Helvetica", pointsize = 12, onefile = TRUE)

VEpbaseline=VEpbasic_waning730f007_1
ymin_value=-1.
# Set margins to reduce space between panels and make room for axis titles
par(mfrow = c(2, 3), mar = c(4.5, 4, 1, 1)+0.5) #+0.5 moves the axes titles closer to the axes
########################### Panel A  ###############################
#######################Different extent of VEs waning##########
plot(c(1:6), VEpbaseline, ylim=c(ymin_value,1),xlab="Study Month", ylab= "VEp", cex=1.5, pch=17, cex.lab=1.5, cex.axis=1.5)
points(c(1:6), VEp4xfaster, col='purple', cex=1.5, pch=16)
points(c(1:6), VEp2xslower, col='blue', cex=1.5, pch=15)
points(c(1:6), VEp4xslower, col='orange', cex=1.5, pch=8)
points(c(1:6), VEpconstantVEs, col='darkcyan', cex=1.5, pch=9)
legend(4,-0.2, # places a legend at the appropriate place 
c("4X faster","baseline", "2X slower","4X slower","VEs=constant"), # puts text in the legend 
pch=c(16,17,15,8,9), # gives the legend appropriate symbols (lines)
bty = "n", #to remove the box around the legend
col=c("purple", "black", "blue","orange","darkcyan"), cex=1.2)
mtext("A", side = 3, line = -1, adj = -0.18, cex = 1.5)
#################
#######################Panel B###################
plot(c(1:6), VEpf001, ylim=c(ymin_value,1),xlab="Study Month", ylab= "VEp", cex=1.5, pch=16, col='purple', cex.lab=1.5, cex.axis=1.5)
points(c(1:6), VEpbaseline, col='black', cex=1.5, pch=17)
points(c(1:6), VEpf015, col='blue', cex=1.5, pch=15)
points(c(1:6), VEpf03, col='orange', cex=1.5, pch=8)
legend(4,-0.2, # places a legend at the appropriate place 
c("f1=f2=0.01","f1=f2=0.07","f1=f2=0.15", "f1=f2=0.3"), # puts text in the legend 
pch=c(16,17,15,8), # gives the legend appropriate symbols (lines)
bty = "n", #to remove the box around the legend
col=c("purple","black","blue","orange"), cex=1.2)
mtext("B", side = 3, line = -1, adj = -0.18, cex = 1.5)
#################
#######################Panel C###################
plot(c(1:6), VEpbaseline, ylim=c(-3.,1),xlab="Study Month", ylab= "VEp", cex=1.5, pch=17, cex.lab=1.5, cex.axis=1.5)
points(c(1:6), VEpbeta2_1, col='purple', cex=1.5, pch=16)
points(c(1:6), VEpbeta2_01, col='blue', cex=1.5, pch=15)
legend(4,-1.8, # places a legend at the appropriate place 
c(expression(paste(beta,"2=0.1")),expression(paste(beta,"2=0.5")),expression(paste(beta,"2=1"))), # puts text in the legend 
pch=c(15,17,16), # gives the legend appropriate symbols (lines)
bty = "n", #to remove the box around the legend
col=c("blue","black", "purple"), cex=1.2)
mtext("C", side = 3, line = -1, adj = -0.18, cex = 1.5)
#################
#######################Panel D###################
plot(c(1:6), VEpbaseline, ylim=c(ymin_value,1),xlab="Study Month", ylab= "VEp", cex=1.5, pch=17, cex.lab=1.5, cex.axis=1.5)
points(c(1:6), VEpalpha4_001, col='blue', cex=1.5, pch=15)
points(c(1:6), VEpalpha4_025, col='brown', cex=1.5, pch=1)
points(c(1:6), VEpalpha4_1, col='purple', cex=1.5, pch=16)
legend(4,-0.2, # places a legend at the appropriate place 
c(expression(paste(alpha,"4=0.01")),expression(paste(alpha,"4=0.03")),expression(paste(alpha,"4=0.25")),expression(paste(alpha,"4=1"))), # puts text in the legend 
pch=c(15,17,1,16), # gives the legend appropriate symbols (lines)
bty = "n", #to remove the box around the legend
col=c("blue", "black", "brown","purple"), cex=1.2)
mtext("D", side = 3, line = -1, adj = -0.18, cex = 1.5)
#################
#######################Panel E###################
plot(c(1:6), VEpbaseline, ylim=c(ymin_value,1),xlab="Study Month", ylab= "VEp", cex=1.5, pch=17, cex.lab=1.5, cex.axis=1.5)
points(c(1:6), VEpbeta4_001, col='purple', cex=1.5, pch=16)
points(c(1:6), VEpbeta4_025, col='blue', cex=1.5, pch=15)
points(c(1:6), VEpbeta4_05, col='orange', cex=1.5, pch=8)
legend(4,-0.2, # places a legend at the appropriate place 
c(expression(paste(beta,"4=0.01")),expression(paste(beta,"4=0.03")),expression(paste(beta,"4=0.25")),expression(paste(beta,"4=0.5"))), # puts text in the legend 
#c("beta4=0.03", "beta4=0.25","beta4=0.5"), # puts text in the legend 
pch=c(16,17,15,8), # gives the legend appropriate symbols (lines)
bty = "n", #to remove the box around the legend
col=c("purple","black", "blue","orange"), cex=1.2)
mtext("E", side = 3, line = -1, adj = -0.18, cex = 1.5)
#################
#######################Panel F###################
#waning of VEp
plot(c(1:6), VEpbaseline, ylim=c(ymin_value,1),xlab="Study Month", ylab= "VEp", cex=1.5, pch=17, cex.lab=1.5, cex.axis=1.5)
points(c(1:6), VEpwaning180, col='darkcyan', cex=1.5, pch=9) #6 month
points(c(1:6), VEpwaning365, col='orange', cex=1.5, pch=8) #1year
points(c(1:6), VEpwaning730, col='blue', cex=1.5, pch=15) #900 days
legend(4,-0.2, # places a legend at the appropriate place 
c("VEp=constant", "2-year waning", "1-year waning","6month waning"), # puts text in the legend 
pch=c(17,15,8,9), # gives the legend appropriate symbols (lines)
bty = "n", #to remove the box around the legend
col=c("black","blue", "orange","darkcyan"), cex=1.2)
mtext("F", side = 3, line = -1, adj = -0.18, cex = 1.5)

dev.off()

knitr::include_graphics("Figure5.pdf")
```

\newpage

**Figure 6. Simulation results from a sensitivity analysis that integrates within-subgroup heterogeneity in VEs and VEp.**

Figure 6 Panels A-C plot the results of sampling individual parameters describing vaccine-induced protection in vulnerable (Group 3) and immunocompetent (Group 4) individuals from corresponding Beta distributions.

```{r figure6_beta_distributions, echo=FALSE, warning=FALSE, message=FALSE}
rm(list=ls()) 

plot_beta_distributions <- function(alpha_immunocompetent, beta_immunocompetent, alpha_vulnerable, beta_vulnerable,
   output_file = "betadistributions.pdf", #save it as default 
   output_file2 = "betadistributions2.pdf", #save it as default 
   vulnerable_color = rgb(0.502, 0, 0.502, 0.8), #rgb(1, 0.65, 0, 0.5), 
   immunocompetent_color = rgb(0, 0.62, 0.451, 0.8)) {   #rgb(0, 0, 1, 0.5)) {
  
  set.seed(123) ## Set seed for reproducibility
  # Generate samples for both groups
  sample_vulnerable <- rbeta(1000, alpha_vulnerable, beta_vulnerable)
  sample_immunocompetent <- rbeta(1000, alpha_immunocompetent, beta_immunocompetent)
  
  # Open PDF to save the plot
  pdf(file = output_file, width = 6.7, height = 5.5, family = "Helvetica", pointsize = 12)
  
  # Adjust plot settings
  par(mgp = c(2.4, 1, 0)) 
  breaks <- seq(0, 1, length.out = 31)  # Define common breaks
  
  # Plot histograms with common breaks
  hist(sample_vulnerable, breaks = breaks, col = vulnerable_color, 
       xlab = "Coefficient", xlim = c(0, 1), ylim = c(0, 800), 
       cex.axis = 1.6, cex.lab = 1.6, main = "")
  hist(sample_immunocompetent, breaks = breaks, col = immunocompetent_color, 
       add = TRUE)
  legend("top", legend = c("Vulnerable", "Immunocompetent"), fill = c(vulnerable_color, immunocompetent_color), cex = 1.5, bty = "n")  

  dev.off() ## Close the PDF device

}

################################ Figure 6 Panel A ###########################################
# Parameters for the two Beta distributions (almost non-overlapping)
#These parameters ensure the variance is approximately 0.01, with the means centered at 0.05 and 0.95, for immunocompetent and vulnerable, respectively.
plot_beta_distributions(0.1875, 3.5625, 3.5625, 0.1875,"Figure6A.pdf")

################################ Figure 6 Panel B ###########################################
#These parameters ensure the variance is approximately 0.01, with the means centered at 0.1 and 0.9, for immunocompetent and vulnerable, respectively.
plot_beta_distributions(0.8, 7.2, 7.2, 0.8,"Figure6B.pdf")

################################ Figure 6 Panel C ###########################################
#These parameters ensure the variance is approximately 0.01, with the means centered at 0.3 and 0.7, for immunocompetent and vulnerable, respectively.
plot_beta_distributions(6, 14, 14, 6,"Figure6C.pdf")

print ("Figure 6A")
knitr::include_graphics("Figure6A.pdf")
print ("Figure 6B")
knitr::include_graphics("Figure6B.pdf")
print ("Figure 6C")
knitr::include_graphics("Figure6C.pdf")
```

Next three chucks generate the data for Figure 6D,E,F. The code is commented, the data files that would be generated  ("VEpbasic_waning730f007_1_0_3_0_7_uncor.csv") are provided for further analysis. 
```{r Figure6F, include=FALSE}

# rm(list=ls()) 
# library(readr)   # Read tabular data.
# library(tidyr)   # Data frame tidying functions.
# library(dplyr)   # General data frame manipulation.
# library(ggplot2) # Flexible plotting.
# 
# VEmodelwithbarplots<-function(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m){ #m is the number of simulations
#   with(as.list(c(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m)),{ # m is the number of the simulation
# # population
# N =200000 #number of unvaccinated people      
# NV=200000 #number of vaccinated people 
# 
# adjust_alpha=1
# adjust_beta=1 #1/0.8 where 0.8 is a mean value for vulnerable group
# 
# #Unvaccinated
# infection_force = 1/180.*adjust_alpha   #4./180. # 30% attack rate over 180 days; to be adjusted
# infection_to_serious_force = 1/10*adjust_beta # base chance to get severe disease if infected
# #Vaccinated
# infection_forceV = 1/180.*adjust_alpha #4./180
# infection_to_serious_forceV = 1/10.*adjust_beta # base chance to get severe disease if infected while being vaccinated
# 
# group1=ceiling(N*frac1)-1 #correction
# group3=ceiling(NV*frac2)-1 #correction
# 
# ####Setting the risk of progression to severe disease######
# # These parameters ensure the variance is approximately 0.01, with the means centered at 0.3 and 0.95, for immunocompetent and vulnerable, respectively.
# alpha_vulnerable <- 14
# beta_vulnerable <- 6
# 
# alpha_immunocompetent <- 6
# beta_immunocompetent <- 14
# 
# sample_vulnerableV <- NULL
# sample_immunocompetentV <- NULL
# 
# # Sample coefficients for both groups for parameter 'beta' in the model
# sample_vulnerableV[1:group3] <- rbeta(group3, alpha_vulnerable, beta_vulnerable)
# sample_immunocompetentV[1:(NV-group3)] <- rbeta((NV-group3), alpha_immunocompetent, beta_immunocompetent)
# 
# sample_vulnerableV_dif <- NULL
# sample_immunocompetentV_dif <- NULL
# 
# #if uncorrelated this is for alpha
# sample_vulnerableV_dif[1:group3] <- rbeta(group3, alpha_vulnerable, beta_vulnerable)
# sample_immunocompetentV_dif[1:(NV-group3)] <- rbeta((NV-group3), alpha_immunocompetent, beta_immunocompetent)
# 
# # #case when correlated
# # sample_vulnerableV_dif[1:group3]=sample_vulnerableV[1:group3]
# # sample_immunocompetentV_dif[1:(NV-group3)]=sample_immunocompetentV[1:(NV-group3)]
# 
# alpha3v=mean(sample_vulnerableV_dif)
# alpha4v=mean(sample_immunocompetentV_dif)
# print(alpha3v)
# print(alpha4v)
# beta3v=mean(sample_vulnerableV)
# beta4v=mean(sample_immunocompetentV)
# print(beta3v)
# print(beta4v)
# 
# if (VEswaning>0) {
# VEs_waining= (alpha2-alpha4v)/VEswaning #(alpha2-alpha4)/180.
# VEs_waining_V= (alpha1-alpha3v)/VEswaning #for vulnerable group
# }
# if (VEpwaning>0) {
# VEp_waining= (beta2-beta4v)/VEpwaning 
# VEp_waining_V= (beta1-beta3v)/VEpwaning 
# }
# 
# ###########################block for unvaccinated##################################
# #initializing groups S and SS with different protection
# S = 0.0*(1:N)
# SS = 0.0*(1:N)
# infect=0*(1:N)
# ser=0*(1:N) #severe cases
# 
# #two groups: 'vulnerable'(group1) and 'immunocompetent'
# S[1:group1]=alpha1 # protection against infection in unvaccinated: 1 means base, <1 means more protected
# S[(group1+1):N]=alpha2
# SS[1:group1]=beta1 # protection against severe disease: 1 means base, <1 means more protected
# SS[(group1+1):N]=beta2 
# 
# for (i in 1:180) #i corresponds to the day, j corresponds to the individal/person
# { # cycle per days
#   for (j in 1:N) # cycle per people
#   {
#     if ((infect[j]==0) && (infection_force*S[j] > runif(1))) # infection probability
#     {
#       infect[j]=i #day of infection
#       if (infection_to_serious_force*SS[j] > runif(1)) # progression to severe probability
#       {
#         ser[j]=i #day of becoming severely ill
#       }
#     }
#   }
# }
# # sum_infect=sum(infect>0)
# # sum_ser= sum(ser>0)
# # ratio_inf_to_ser = sum_infect/sum_ser
# 
# #calculate at what month each person was infected; 0 - person is not infected
# infect_month=(infect+30)%/%31 
# 
# a=data.frame(S,infect,infect_month,ser)
# b = a %>% group_by(infect) %>% summarise(mean=mean(S))
# 
# #infect_month=0 in 'c' shows the line for uninfected; if absent -all infected; if '6' is absent - noone infected during month 6
# c = a %>% group_by(infect_month) %>% summarise(mean_per_Month=mean(S),N=n())
# 
# ser_month=(ser+30)%/%30.1
# a_ss=data.frame(S,ser,ser_month)
# ##ser_month=0 in 'd' shows the line for those not progressed to severe disease 
# d = a_ss %>% group_by(ser_month) %>% summarise(SS_per_Month=mean(S),Nser=n())
# 
# ###############################################vaccinated##############################
# #initializing groups S and SS with different protection
# SV = 0.0*(1:NV)
# SSV = 0.0*(1:NV)
# infectV=0*(1:NV)
# serV=0*(1:NV)
# 
# #two groups: 'vulnerable'(group3) and 'immunocompetent'(group 4) in vaccinated
# 
# if (VEswaning==0) {
#   SV[1:group3]=sample_vulnerableV_dif[1:group3] 
#   SV[(group3+1):NV]=sample_immunocompetentV_dif[1:(NV-group3)]
# }
# 
# if (VEpwaning==0) {
# # #two lines below are needed only for stable progression to severe disease#
# # #also need to comment the line SSV=SV below
#   SSV[1:group3]=sample_vulnerableV[1:group3]
#   SSV[(group3+1):NV]=sample_immunocompetentV[1:(NV-group3)]  #beta4
# }
# 
# for (i in 1:180)  { # cycle per days
#   if (VEswaning>0) {
#   SV[1:group3]=sample_vulnerableV_dif[1:group3] #alpha3
#   SV[(group3+1):NV]=sample_immunocompetentV_dif[1:(NV-group3)] #alpha4
#   if(mean(sample_immunocompetentV_dif[1:(NV-group3)])<alpha2){
#   sample_immunocompetentV_dif[1:(NV-group3)]= sample_immunocompetentV_dif[1:(NV-group3)]+VEs_waining
#  for (schet in 1:(NV-group3)){
#     if (sample_immunocompetentV_dif[schet]>1) {sample_immunocompetentV_dif[schet]==1}
#   }
#   }
#   if(mean(sample_vulnerableV_dif[1:group3])<alpha1){
#   sample_vulnerableV_dif[1:group3]= sample_vulnerableV_dif[1:group3]+VEs_waining_V
#     for (schet in 1:group3){
#     if (sample_vulnerableV_dif[schet]>1) {sample_vulnerableV_dif[schet]==1}
#   }
#   }
#   }
#   if (VEpwaning>0) {
#   SSV[(group3+1):NV]=sample_immunocompetentV[1:(NV-group3)]  #beta4
#   SSV[1:group3]=sample_vulnerableV[1:group3]  #beta3
#   if(mean(sample_immunocompetentV[1:(NV-group3)])<beta2){
#   sample_immunocompetentV[1:(NV-group3)]= sample_immunocompetentV[1:(NV-group3)]+VEp_waining
#   for (schet in 1:(NV-group3)){
#     if (sample_immunocompetentV[schet]>1) {sample_immunocompetentV[schet]==1}
#   }
#   }
#   if(mean(sample_vulnerableV[1:group3])<beta1){
#   sample_vulnerableV[1:group3]= sample_vulnerableV[1:group3]+VEp_waining_V
#   for (schet in 1:group3){
#     if (sample_vulnerableV[schet]>1) {sample_vulnerableV[schet]==1}
#   }
#   }
#   }
# 
#   for (j in 1:N) # cycle per people
#   {
#     if ((infectV[j]==0) && (infection_forceV*SV[j] > runif(1))) # infection probability
#     {
#       infectV[j]=i
#       if (infection_to_serious_forceV*SSV[j] > runif(1)) # serious probability
#       {
#         serV[j]=i
#       }
#     }
#   }
# }
# sum_infectV=sum(infectV>0)
# sum_serV= sum(serV>0)
# ratio_inf_to_serV = sum(infectV>0)/sum(serV>0)
# 
# infect_monthV=(infectV+30)%/%31
# aV=data.frame(SV,infectV,infect_monthV,serV)
# bV = aV %>% group_by(infectV) %>% summarise(mean=mean(SV))
# cV = aV %>% group_by(infect_monthV) %>% summarise(mean_per_MonthV=mean(SV),NV=n())
#  
# ser_monthV=(serV+30)%/%30.1
# a_ssV=data.frame(SV,serV,ser_monthV)
# dV = a_ssV %>% group_by(ser_monthV) %>% summarise(SS_per_MonthV=mean(SV),NserV=n())
# 
# #########generate file to report susceptible left######
# susceptible=data.frame(S,infect,infect_month,ser,SV,infectV,infect_monthV,serV)
# write.csv(susceptible, file = paste0("susceptible1_0_3_0_7_uncor_",m,".csv"), row.names = FALSE)
# ###############################################################################################
# ######################################BAR PLOTS################################################
# # adding labels
# SG = rep('', times=N) # group label
# SG[1:group1]='1' # 
# SG[(group1+1):N]='2'
# 
# a_per_group=data.frame(SG,S,infect_month)
# c_per_group = a_per_group %>%
#   filter(infect_month > 0 ) %>%
#   group_by(SG,infect_month) %>% 
#   summarise(mean_per_Month=mean(S),N_inf=n())
# # plotting
# barplotU = data.frame(G = c_per_group$SG ,Month = c_per_group$infect_month,N_inf = c_per_group$N_inf)
# 
# SVG = rep('', times=N) #group label
# SVG[1:group3]="3"
# SVG[(group3+1):NV]="4"
# 
# ##############################Bar Plot for Infected##################################
# av_per_group=data.frame(SVG,SV,infect_monthV)
# cv_per_group = av_per_group %>%
#   filter(infect_monthV > 0 ) %>%
#   group_by(SVG,infect_monthV) %>% 
#   summarise(mean_per_Month=mean(SV),N_inf=n())
# barplotV = data.frame(G = cv_per_group$SVG ,Month = cv_per_group$infect_monthV,N_inf= cv_per_group$N_inf)
# 
# # all four groups combined for the bar plot
# barplotALL=rbind(barplotU,barplotV)
# #ggplot(barplotALL, aes(x = Month, y = N_inf, fill = G, colour = G)) + geom_bar(stat = "identity")
# #ggsave("barplotALL.eps", width = 6, height = 4.5, device="eps") 
# write.csv(barplotALL, file = paste0("barplotALL1_0_3_0_7_uncor_",m,".csv"), row.names = FALSE)
# 
# ##############################Bar Plot for Severe##################################
# #################progressed to severe in unvaccinated###################
# as_per_group=data.frame(SG,S,ser_month)
# cs_per_group = as_per_group %>%
#   filter(ser_month > 0 ) %>%
#   group_by(SG,ser_month) %>% 
#   summarise(mean_per_Month=mean(S),N_ser=n())
# barplotSU = data.frame(G = cs_per_group$SG ,Month = cs_per_group$ser_month,N_ser= cs_per_group$N_ser)
# #####progressed to severe in vaccinated###################
# avs_per_group=data.frame(SVG,SV,ser_monthV)
# cvs_per_group = avs_per_group %>%
#   filter(ser_monthV > 0 ) %>%
#   group_by(SVG,ser_monthV) %>% 
#   summarise(mean_per_Month=mean(SV),N_ser=n())
# barplotSV = data.frame(G = cvs_per_group$SVG ,Month = cvs_per_group$ser_monthV,N_ser= cvs_per_group$N_ser)
# 
# barplotALLS=rbind(barplotSU,barplotSV)
# write.csv(barplotALLS, file = paste0("barplotALLS1_0_3_0_7_uncor_",m,".csv"), row.names = FALSE)
# ###############################################################################################
# ##################################calculating VEp###############################
# VEp=1-(dV$NserV/cV$NV)[2:7]/(d$Nser/c$N)[2:7] 
#    return(VEp)
#    })
# }
# set.seed(123) #setting seed keeps ALL 100 simulations the same if you repeat it
# simul_number=100
# VEpbasic <- matrix(NA,nrow=simul_number,ncol=6)
# for (m in 1:simul_number) {
# VEpbasic[m,] <- VEmodelwithbarplots(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,730,0,m)
# }
# write.csv(VEpbasic, file = "VEpbasic_waning730f007_1_0_3_0_7_uncor.csv", row.names = FALSE)

```

```{r figure6DEFplots, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}

###########################  Figure 6 Panels D, E, and F  #################################
#Finding the closest to the average curve and it's number

########Plotting all 100 simulation and the one that is closest to average####################

Plotting100<-function(filename){ 
  with(as.list(c(filename)),{
    
# Read the CSV file
file = paste0(filename, ".csv")
data <- read.csv(file)
  
strings <- data
# Calculate the overall average values 
overall_averaged_values <- apply(strings, 2, mean) 

# Function to find the string with the closest values to the overall averaged values 
find_closest_string <- function(strings, overall_averaged_values) { 
  # Calculate the distances between each string and the overall averaged values 
  distances <- apply(strings, 1, function(x) sqrt(sum((x - overall_averaged_values)^2))) 
  # Find the index of the string with the minimum distance 
  closest_string_index <- which.min(distances) 
  return(list(string = strings[closest_string_index, ], index = closest_string_index)) 
  } 
# Find the closest string to the overall averaged values 
closest_string <- find_closest_string(strings, overall_averaged_values) 

pdf(file = paste0(filename, ".pdf"), width = 6.7, height = 5.5, family = "Helvetica", pointsize = 12, onefile = TRUE)
# Plot all strings 
matplot(1:6, t(strings), type = "l", col = "grey", lty = 1, xlab = "Study Month", ylab = "VEp",ylim=c(0,0.8)) 
# Plot the chosen closest string as a thick line 
lines(1:6, closest_string$string, type = "l", lwd = 2) 
#legend("topleft", legend = paste("Closest to Average (VEp#", closest_string$index,")", sep = " "), col = "black", lwd = 2)
dev.off()
     return(closest_string$index)
   })
 }

################finding the closest to averaged VE values for each condition ############ 
allfilenames <- c("VEpbasic_waning730f007_0_05_0_95", "VEpbasic_waning730f007_heter0_1_0_9","VEpbasic_waning730f007_1_0_3_0_7_uncor")

for (k in 1:length(allfilenames)) {
index=Plotting100(allfilenames[k])
file = paste0(allfilenames[k], ".csv")
data=read.csv(file)

# Extract the specific row and convert it to a numeric vector
# Create the variable with the name in the name vector
#the result is a varibale with the name allfilenames[i] that has the simulation values closest to the 100 simulation average
assign(allfilenames[k], as.numeric(data[index,]), envir = .GlobalEnv)
}

print("Figure 6D")
knitr::include_graphics("VEpbasic_waning730f007_0_05_0_95.pdf")
print("Figure 6E")
knitr::include_graphics("VEpbasic_waning730f007_heter0_1_0_9.pdf")
print("Figure 6F")
knitr::include_graphics("VEpbasic_waning730f007_1_0_3_0_7_uncor.pdf")
```

\newpage

**Figure 7. Effect of high vaccineinduced protection in immunocompetent group and low vaccineinduced protection in vulnerable group on observed VEp.** 

Figure 7 Panels A-C plot the results of sampling individual parameters describing vaccine-induced protection in vulnerable (Group 3) and immunocompetent (Group 4) individuals from corresponding Beta distributions.

```{r figure7_beta_distributions, echo=FALSE, warning=FALSE, message=FALSE}
rm(list=ls()) 

plot_beta_distributions <- function(alpha_immunocompetent, beta_immunocompetent, alpha_vulnerable, beta_vulnerable,
   output_file = "betadistributions.pdf", #save it as default 
   output_file2 = "betadistributions2.pdf", #save it as default 
   vulnerable_color = rgb(0.502, 0, 0.502, 0.8), #rgb(1, 0.65, 0, 0.5), 
   immunocompetent_color = rgb(0, 0.62, 0.451, 0.8)) {   #rgb(0, 0, 1, 0.5)) {
  
  set.seed(123) ## Set seed for reproducibility
  # Generate samples for both groups
  sample_vulnerable <- rbeta(1000, alpha_vulnerable, beta_vulnerable)
  sample_immunocompetent <- rbeta(1000, alpha_immunocompetent, beta_immunocompetent)
  
  # Open PDF to save the plot
  pdf(file = output_file, width = 6.7, height = 5.5, family = "Helvetica", pointsize = 12)
  
  # Adjust plot settings
  par(mgp = c(2.4, 1, 0)) 
  breaks <- seq(0, 1, length.out = 31)  # Define common breaks
  
  # Plot histograms with common breaks
  hist(sample_vulnerable, breaks = breaks, col = vulnerable_color, 
       xlab = "Coefficient", xlim = c(0, 1), ylim = c(0, 800), 
       cex.axis = 1.6, cex.lab = 1.6, main = "")
  hist(sample_immunocompetent, breaks = breaks, col = immunocompetent_color, 
       add = TRUE)
  legend("top", legend = c("Vulnerable", "Immunocompetent"), fill = c(vulnerable_color, immunocompetent_color), cex = 1.5, bty = "n")  

  dev.off() ## Close the PDF device

}

############################################ Figure 7 Panel A ###########################################
# Parameters for the two Beta distributions (almost non-overlapping)
#These parameters ensure the variance is approximately 0.01, with the means centered at 0.05 and 0.95, for immunocompetent and vulnerable, respectively.
plot_beta_distributions(0.1875, 3.5625, 3.5625, 0.1875,"Figure7A.pdf")

################################## Figure 7 Panel B #########################################
#These parameters ensure the variance is approximately 0.01, with the means centered at 0.05 and 0.7, for immunocompetent and vulnerable, respectively.
plot_beta_distributions(0.1875, 3.5625, 14, 6,"Figure7B.pdf")

################################## Figure 7 Panel C ########################################
#means 0.3 and 0.95 and variance 0.01
plot_beta_distributions(6, 14, 3.5625, 0.1875,"Figure7C.pdf")

knitr::include_graphics("Figure7A.pdf")
knitr::include_graphics("Figure7B.pdf")
knitr::include_graphics("Figure7C.pdf")
```

Figure 7D is the same as Figure 6D. Next two chucks generate the data for Figure 7E and Figure 7F, following by the chunk that plots the results.  
The files "VEpbasic_waning730f007uncor0_05_0_7.csv" (Panel E) and "VEpbasic_waning730f007_1_0_3_0_95_uncor.csv" (Panel F) are provided for further analysis.

```{r figure7panelE, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}

########Uncorrelated  means 0.7 (vulnerable) and 0.05 (immunocompetent) and a variance of 0.01###

rm(list=ls()) 
library(readr)   # Read tabular data.
library(tidyr)   # Data frame tidying functions.
library(dplyr)   # General data frame manipulation.
library(ggplot2) # Flexible plotting.

# VEmodelwithbarplots<-function(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m){ #m is the number of simulations
#   with(as.list(c(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m)),{ # m is the number of the simulation
# # population
# N =200000 #number of unvaccinated people      
# NV=200000 #number of vaccinated people 
# 
# adjust_alpha=1
# adjust_beta=1 #1/0.8 where 0.8 is a mean value for vulnerable group
# 
# #Unvaccinated
# infection_force = 1/180.*adjust_alpha   #4./180. # 30% attack rate over 180 days; to be adjusted
# infection_to_serious_force = 1/10*adjust_beta # base chance to get severe disease if infected
# #Vaccinated
# infection_forceV = 1/180.*adjust_alpha #4./180
# infection_to_serious_forceV = 1/10.*adjust_beta # base chance to get severe disease if infected while being vaccinated
# 
# group1=ceiling(N*frac1)-1 #correction
# group3=ceiling(NV*frac2)-1 #correction
# 
# sample_vulnerableV <- NULL
# sample_immunocompetentV <- NULL
# 
# ######################################## Panel E ##################################
# ##### Beta ditribution means are 0.7 (vulnerable) and 0.05 (immunocompetent) and a variance of 0.01#####
# alpha_vulnerable <- 14
# beta_vulnerable <- 6
# 
# alpha_immunocompetent <- 0.1875
# beta_immunocompetent <- 3.5625

# # Sample coefficients for both groups for parameter 'beta' in the model
# sample_vulnerableV[1:group3] <- rbeta(group3, alpha_vulnerable, beta_vulnerable)
# sample_immunocompetentV[1:(NV-group3)] <- rbeta((NV-group3), alpha_immunocompetent, beta_immunocompetent)
# 
# #if uncorrelated this is for alpha
# sample_vulnerableV_dif <- NULL
# sample_immunocompetentV_dif <- NULL
# 
# #case when individual VEs and VEp areuncorrelated
# sample_vulnerableV_dif[1:group3] <- rbeta(group3, alpha_vulnerable, beta_vulnerable)
# sample_immunocompetentV_dif[1:(NV-group3)] <- rbeta((NV-group3), alpha_immunocompetent, beta_immunocompetent)
# 
# # #case when individual VEs and VEp are correlated as in Supplemental Figure 2
# # sample_vulnerableV_dif[1:group3]=sample_vulnerableV[1:group3]
# # sample_immunocompetentV_dif[1:(NV-group3)]=sample_immunocompetentV[1:(NV-group3)]
# 
# alpha3v=mean(sample_vulnerableV_dif)
# alpha4v=mean(sample_immunocompetentV_dif)
# print(alpha3v)
# print(alpha4v)
# beta3v=mean(sample_vulnerableV)
# beta4v=mean(sample_immunocompetentV)
# print(beta3v)
# print(beta4v)
# 
# if (VEswaning>0) {
# VEs_waining= (alpha2-alpha4v)/VEswaning #(alpha2-alpha4)/180.
# VEs_waining_V= (alpha1-alpha3v)/VEswaning #for vulnerable group
# }
# if (VEpwaning>0) {
# VEp_waining= (beta2-beta4v)/VEpwaning 
# VEp_waining_V= (beta1-beta3v)/VEpwaning 
# }
# 
# ###########################block for unvaccinated##################################
# #initializing groups S and SS with different protection
# S = 0.0*(1:N)
# SS = 0.0*(1:N)
# infect=0*(1:N)
# ser=0*(1:N) #severe cases
# 
# #two groups: 'vulnerable'(group1) and 'immunocompetent'
# S[1:group1]=alpha1 # protection against infection in unvaccinated: 1 means base, <1 means more protected
# S[(group1+1):N]=alpha2
# SS[1:group1]=beta1 # protection against severe disease: 1 means base, <1 means more protected
# SS[(group1+1):N]=beta2 
# 
# for (i in 1:180) #i corresponds to the day, j corresponds to the individal/person
# { # cycle per days
#   for (j in 1:N) # cycle per people
#   {
#     if ((infect[j]==0) && (infection_force*S[j] > runif(1))) # infection probability
#     {
#       infect[j]=i #day of infection
#       if (infection_to_serious_force*SS[j] > runif(1)) # progression to severe probability
#       {
#         ser[j]=i #day of becoming severely ill
#       }
#     }
#   }
# }
# # sum_infect=sum(infect>0)
# # sum_ser= sum(ser>0)
# # ratio_inf_to_ser = sum_infect/sum_ser
# 
# #calculate at what month each person was infected; 0 - person is not infected
# infect_month=(infect+30)%/%31 
# 
# a=data.frame(S,infect,infect_month,ser)
# b = a %>% group_by(infect) %>% summarise(mean=mean(S))
# 
# #infect_month=0 in 'c' shows the line for uninfected; if absent -all infected; if '6' is absent - noone infected during month 6
# c = a %>% group_by(infect_month) %>% summarise(mean_per_Month=mean(S),N=n())
# 
# ser_month=(ser+30)%/%30.1
# a_ss=data.frame(S,ser,ser_month)
# ##ser_month=0 in 'd' shows the line for those not progressed to severe disease 
# d = a_ss %>% group_by(ser_month) %>% summarise(SS_per_Month=mean(S),Nser=n())
# 
# ###############################################vaccinated##############################
# #initializing groups S and SS with different protection
# SV = 0.0*(1:NV)
# SSV = 0.0*(1:NV)
# infectV=0*(1:NV)
# serV=0*(1:NV)
# 
# #two groups: 'vulnerable'(group3) and 'immunocompetent'(group 4) in vaccinated
# 
# if (VEswaning==0) {
#   SV[1:group3]=sample_vulnerableV_dif[1:group3] 
#   SV[(group3+1):NV]=sample_immunocompetentV_dif[1:(NV-group3)]
# }
# 
# if (VEpwaning==0) {
# # #two lines below are needed only for stable progression to severe disease#
# # #also need to comment the line SSV=SV below
#   SSV[1:group3]=sample_vulnerableV[1:group3]
#   SSV[(group3+1):NV]=sample_immunocompetentV[1:(NV-group3)]  #beta4
# }
# 
# for (i in 1:180)  { # cycle per days
#   if (VEswaning>0) {
#   SV[1:group3]=sample_vulnerableV_dif[1:group3] #alpha3
#   SV[(group3+1):NV]=sample_immunocompetentV_dif[1:(NV-group3)] #alpha4
#   if(mean(sample_immunocompetentV_dif[1:(NV-group3)])<alpha2){
#   sample_immunocompetentV_dif[1:(NV-group3)]= sample_immunocompetentV_dif[1:(NV-group3)]+VEs_waining
#  for (schet in 1:(NV-group3)){
#     if (sample_immunocompetentV_dif[schet]>1) {sample_immunocompetentV_dif[schet]==1}
#   }
#   }
#   if(mean(sample_vulnerableV_dif[1:group3])<alpha1){
#   sample_vulnerableV_dif[1:group3]= sample_vulnerableV_dif[1:group3]+VEs_waining_V
#     for (schet in 1:group3){
#     if (sample_vulnerableV_dif[schet]>1) {sample_vulnerableV_dif[schet]==1}
#   }
#   }
#   }
#   if (VEpwaning>0) {
#   SSV[(group3+1):NV]=sample_immunocompetentV[1:(NV-group3)]  #beta4
#   SSV[1:group3]=sample_vulnerableV[1:group3]  #beta3
#   if(mean(sample_immunocompetentV[1:(NV-group3)])<beta2){
#   sample_immunocompetentV[1:(NV-group3)]= sample_immunocompetentV[1:(NV-group3)]+VEp_waining
#   for (schet in 1:(NV-group3)){
#     if (sample_immunocompetentV[schet]>1) {sample_immunocompetentV[schet]==1}
#   }
#   }
#   if(mean(sample_vulnerableV[1:group3])<beta1){
#   sample_vulnerableV[1:group3]= sample_vulnerableV[1:group3]+VEp_waining_V
#   for (schet in 1:group3){
#     if (sample_vulnerableV[schet]>1) {sample_vulnerableV[schet]==1}
#   }
#   }
#   }
# 
#   for (j in 1:N) # cycle per people
#   {
#     if ((infectV[j]==0) && (infection_forceV*SV[j] > runif(1))) # infection probability
#     {
#       infectV[j]=i
#       if (infection_to_serious_forceV*SSV[j] > runif(1)) # serious probability
#       {
#         serV[j]=i
#       }
#     }
#   }
# }
# sum_infectV=sum(infectV>0)
# sum_serV= sum(serV>0)
# ratio_inf_to_serV = sum(infectV>0)/sum(serV>0)
# 
# infect_monthV=(infectV+30)%/%31
# aV=data.frame(SV,infectV,infect_monthV,serV)
# bV = aV %>% group_by(infectV) %>% summarise(mean=mean(SV))
# cV = aV %>% group_by(infect_monthV) %>% summarise(mean_per_MonthV=mean(SV),NV=n())
#  
# ser_monthV=(serV+30)%/%30.1
# a_ssV=data.frame(SV,serV,ser_monthV)
# dV = a_ssV %>% group_by(ser_monthV) %>% summarise(SS_per_MonthV=mean(SV),NserV=n())
# 
# #########generate file to report susceptible left######
# susceptible=data.frame(S,infect,infect_month,ser,SV,infectV,infect_monthV,serV)
# write.csv(susceptible, file = paste0("susceptible1_uncor0_05_0_7_",m,".csv"), row.names = FALSE)
# ###############################################################################################
# ######################################BAR PLOTS################################################
# # adding labels
# SG = rep('', times=N) # group label
# SG[1:group1]='1' # 
# SG[(group1+1):N]='2'
# 
# a_per_group=data.frame(SG,S,infect_month)
# c_per_group = a_per_group %>%
#   filter(infect_month > 0 ) %>%
#   group_by(SG,infect_month) %>% 
#   summarise(mean_per_Month=mean(S),N_inf=n())
# # plotting
# barplotU = data.frame(G = c_per_group$SG ,Month = c_per_group$infect_month,N_inf = c_per_group$N_inf)
# 
# SVG = rep('', times=N) #group label
# SVG[1:group3]="3"
# SVG[(group3+1):NV]="4"
# 
# ##############################Bar Plot for Infected##################################
# av_per_group=data.frame(SVG,SV,infect_monthV)
# cv_per_group = av_per_group %>%
#   filter(infect_monthV > 0 ) %>%
#   group_by(SVG,infect_monthV) %>% 
#   summarise(mean_per_Month=mean(SV),N_inf=n())
# barplotV = data.frame(G = cv_per_group$SVG ,Month = cv_per_group$infect_monthV,N_inf= cv_per_group$N_inf)
# 
# # all four groups combined for the bar plot
# barplotALL=rbind(barplotU,barplotV)
# #ggplot(barplotALL, aes(x = Month, y = N_inf, fill = G, colour = G)) + geom_bar(stat = "identity")
# #ggsave("barplotALL.eps", width = 6, height = 4.5, device="eps") 
# write.csv(barplotALL, file = paste0("barplotALLuncor0_05_0_7_",m,".csv"), row.names = FALSE)
# 
# ##############################Bar Plot for Severe##################################
# #################progressed to severe in unvaccinated###################
# as_per_group=data.frame(SG,S,ser_month)
# cs_per_group = as_per_group %>%
#   filter(ser_month > 0 ) %>%
#   group_by(SG,ser_month) %>% 
#   summarise(mean_per_Month=mean(S),N_ser=n())
# barplotSU = data.frame(G = cs_per_group$SG ,Month = cs_per_group$ser_month,N_ser= cs_per_group$N_ser)
# #####progressed to severe in vaccinated###################
# avs_per_group=data.frame(SVG,SV,ser_monthV)
# cvs_per_group = avs_per_group %>%
#   filter(ser_monthV > 0 ) %>%
#   group_by(SVG,ser_monthV) %>% 
#   summarise(mean_per_Month=mean(SV),N_ser=n())
# barplotSV = data.frame(G = cvs_per_group$SVG ,Month = cvs_per_group$ser_monthV,N_ser= cvs_per_group$N_ser)
# 
# barplotALLS=rbind(barplotSU,barplotSV)
# write.csv(barplotALLS, file = paste0("barplotALLSuncor0_05_0_7_",m,".csv"), row.names = FALSE)
# ###############################################################################################
# ##################################calculating VEp###############################
# VEp=1-(dV$NserV/cV$NV)[2:7]/(d$Nser/c$N)[2:7] 
#    return(VEp)
#    })
# }
# 
# set.seed(123) #setting seed keeps ALL 100 simulations the same if you repeat it
# simul_number=100
# VEpbasic <- matrix(NA,nrow=simul_number,ncol=6)
# for (m in 1:simul_number) {
# VEpbasic[m,] <- VEmodelwithbarplots(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,730,0,m)
# }
# write.csv(VEpbasic, file = "VEpbasic_waning730f007uncor0_05_0_7.csv", row.names = FALSE)

############Uncomment the lines above to generate VEpbasic_waning730f007uncor0_05_0_7.csv ######
```


```{r figure7F, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}

# rm(list=ls()) 
# library(readr)   # Read tabular data.
# library(tidyr)   # Data frame tidying functions.
# library(dplyr)   # General data frame manipulation.
# library(ggplot2) # Flexible plotting.
# 
# VEmodelwithbarplots<-function(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m){ #m is the number of simulations
#   with(as.list(c(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m)),{ # m is the number of the simulation
# # population
# N =200000 #number of unvaccinated people      
# NV=200000 #number of vaccinated people 
# 
# adjust_alpha=1
# adjust_beta=1 #1/0.8 where 0.8 is a mean value for vulnerable group
# 
# #Unvaccinated
# infection_force = 1/180.*adjust_alpha   #4./180. # 30% attack rate over 180 days; to be adjusted
# infection_to_serious_force = 1/10*adjust_beta # base chance to get severe disease if infected
# #Vaccinated
# infection_forceV = 1/180.*adjust_alpha #4./180
# infection_to_serious_forceV = 1/10.*adjust_beta # base chance to get severe disease if infected while being vaccinated
# 
# group1=ceiling(N*frac1)-1 #correction
# group3=ceiling(NV*frac2)-1 #correction
# 
# ####Setting the risk of progression to severe disease######
# # These parameters ensure the variance is approximately 0.01, with the means centered at 0.3 and 0.95, for immunocompetent and vulnerable, respectively.
# alpha_vulnerable <- 3.5625
# beta_vulnerable <- 0.1875
# 
# alpha_immunocompetent <- 6
# beta_immunocompetent <- 14
# 
# sample_vulnerableV <- NULL
# sample_immunocompetentV <- NULL
# 
# # Sample coefficients for both groups for parameter 'beta' in the model
# sample_vulnerableV[1:group3] <- rbeta(group3, alpha_vulnerable, beta_vulnerable)
# sample_immunocompetentV[1:(NV-group3)] <- rbeta((NV-group3), alpha_immunocompetent, beta_immunocompetent)
# 
# sample_vulnerableV_dif <- NULL
# sample_immunocompetentV_dif <- NULL
# 
# #if uncorrelated this is for alpha
# sample_vulnerableV_dif[1:group3] <- rbeta(group3, alpha_vulnerable, beta_vulnerable)
# sample_immunocompetentV_dif[1:(NV-group3)] <- rbeta((NV-group3), alpha_immunocompetent, beta_immunocompetent)
# 
# # #case when correlated
# # sample_vulnerableV_dif[1:group3]=sample_vulnerableV[1:group3]
# # sample_immunocompetentV_dif[1:(NV-group3)]=sample_immunocompetentV[1:(NV-group3)]
# 
# alpha3v=mean(sample_vulnerableV_dif)
# alpha4v=mean(sample_immunocompetentV_dif)
# print(alpha3v)
# print(alpha4v)
# beta3v=mean(sample_vulnerableV)
# beta4v=mean(sample_immunocompetentV)
# print(beta3v)
# print(beta4v)
# 
# if (VEswaning>0) {
# VEs_waining= (alpha2-alpha4v)/VEswaning #(alpha2-alpha4)/180.
# VEs_waining_V= (alpha1-alpha3v)/VEswaning #for vulnerable group
# }
# if (VEpwaning>0) {
# VEp_waining= (beta2-beta4v)/VEpwaning 
# VEp_waining_V= (beta1-beta3v)/VEpwaning 
# }
# 
# ###########################block for unvaccinated##################################
# #initializing groups S and SS with different protection
# S = 0.0*(1:N)
# SS = 0.0*(1:N)
# infect=0*(1:N)
# ser=0*(1:N) #severe cases
# 
# #two groups: 'vulnerable'(group1) and 'immunocompetent'
# S[1:group1]=alpha1 # protection against infection in unvaccinated: 1 means base, <1 means more protected
# S[(group1+1):N]=alpha2
# SS[1:group1]=beta1 # protection against severe disease: 1 means base, <1 means more protected
# SS[(group1+1):N]=beta2 
# 
# for (i in 1:180) #i corresponds to the day, j corresponds to the individal/person
# { # cycle per days
#   for (j in 1:N) # cycle per people
#   {
#     if ((infect[j]==0) && (infection_force*S[j] > runif(1))) # infection probability
#     {
#       infect[j]=i #day of infection
#       if (infection_to_serious_force*SS[j] > runif(1)) # progression to severe probability
#       {
#         ser[j]=i #day of becoming severely ill
#       }
#     }
#   }
# }
# # sum_infect=sum(infect>0)
# # sum_ser= sum(ser>0)
# # ratio_inf_to_ser = sum_infect/sum_ser
# 
# #calculate at what month each person was infected; 0 - person is not infected
# infect_month=(infect+30)%/%31 
# 
# a=data.frame(S,infect,infect_month,ser)
# b = a %>% group_by(infect) %>% summarise(mean=mean(S))
# 
# #infect_month=0 in 'c' shows the line for uninfected; if absent -all infected; if '6' is absent - noone infected during month 6
# c = a %>% group_by(infect_month) %>% summarise(mean_per_Month=mean(S),N=n())
# 
# ser_month=(ser+30)%/%30.1
# a_ss=data.frame(S,ser,ser_month)
# ##ser_month=0 in 'd' shows the line for those not progressed to severe disease 
# d = a_ss %>% group_by(ser_month) %>% summarise(SS_per_Month=mean(S),Nser=n())
# 
# ###############################################vaccinated##############################
# #initializing groups S and SS with different protection
# SV = 0.0*(1:NV)
# SSV = 0.0*(1:NV)
# infectV=0*(1:NV)
# serV=0*(1:NV)
# 
# #two groups: 'vulnerable'(group3) and 'immunocompetent'(group 4) in vaccinated
# 
# if (VEswaning==0) {
#   SV[1:group3]=sample_vulnerableV_dif[1:group3] 
#   SV[(group3+1):NV]=sample_immunocompetentV_dif[1:(NV-group3)]
# }
# 
# if (VEpwaning==0) {
# # #two lines below are needed only for stable progression to severe disease#
# # #also need to comment the line SSV=SV below
#   SSV[1:group3]=sample_vulnerableV[1:group3]
#   SSV[(group3+1):NV]=sample_immunocompetentV[1:(NV-group3)]  #beta4
# }
# 
# for (i in 1:180)  { # cycle per days
#   if (VEswaning>0) {
#   SV[1:group3]=sample_vulnerableV_dif[1:group3] #alpha3
#   SV[(group3+1):NV]=sample_immunocompetentV_dif[1:(NV-group3)] #alpha4
#   if(mean(sample_immunocompetentV_dif[1:(NV-group3)])<alpha2){
#   sample_immunocompetentV_dif[1:(NV-group3)]= sample_immunocompetentV_dif[1:(NV-group3)]+VEs_waining
#  for (schet in 1:(NV-group3)){
#     if (sample_immunocompetentV_dif[schet]>1) {sample_immunocompetentV_dif[schet]==1}
#   }
#   }
#   if(mean(sample_vulnerableV_dif[1:group3])<alpha1){
#   sample_vulnerableV_dif[1:group3]= sample_vulnerableV_dif[1:group3]+VEs_waining_V
#     for (schet in 1:group3){
#     if (sample_vulnerableV_dif[schet]>1) {sample_vulnerableV_dif[schet]==1}
#   }
#   }
#   }
#   if (VEpwaning>0) {
#   SSV[(group3+1):NV]=sample_immunocompetentV[1:(NV-group3)]  #beta4
#   SSV[1:group3]=sample_vulnerableV[1:group3]  #beta3
#   if(mean(sample_immunocompetentV[1:(NV-group3)])<beta2){
#   sample_immunocompetentV[1:(NV-group3)]= sample_immunocompetentV[1:(NV-group3)]+VEp_waining
#   for (schet in 1:(NV-group3)){
#     if (sample_immunocompetentV[schet]>1) {sample_immunocompetentV[schet]==1}
#   }
#   }
#   if(mean(sample_vulnerableV[1:group3])<beta1){
#   sample_vulnerableV[1:group3]= sample_vulnerableV[1:group3]+VEp_waining_V
#   for (schet in 1:group3){
#     if (sample_vulnerableV[schet]>1) {sample_vulnerableV[schet]==1}
#   }
#   }
#   }
# 
#   for (j in 1:N) # cycle per people
#   {
#     if ((infectV[j]==0) && (infection_forceV*SV[j] > runif(1))) # infection probability
#     {
#       infectV[j]=i
#       if (infection_to_serious_forceV*SSV[j] > runif(1)) # serious probability
#       {
#         serV[j]=i
#       }
#     }
#   }
# }
# sum_infectV=sum(infectV>0)
# sum_serV= sum(serV>0)
# ratio_inf_to_serV = sum(infectV>0)/sum(serV>0)
# 
# infect_monthV=(infectV+30)%/%31
# aV=data.frame(SV,infectV,infect_monthV,serV)
# bV = aV %>% group_by(infectV) %>% summarise(mean=mean(SV))
# cV = aV %>% group_by(infect_monthV) %>% summarise(mean_per_MonthV=mean(SV),NV=n())
#  
# ser_monthV=(serV+30)%/%30.1
# a_ssV=data.frame(SV,serV,ser_monthV)
# dV = a_ssV %>% group_by(ser_monthV) %>% summarise(SS_per_MonthV=mean(SV),NserV=n())
# 
# #########generate file to report susceptible left######
# susceptible=data.frame(S,infect,infect_month,ser,SV,infectV,infect_monthV,serV)
# write.csv(susceptible, file = paste0("susceptible1_0_3_0_95_uncor_",m,".csv"), row.names = FALSE)
# ###############################################################################################
# ######################################BAR PLOTS################################################
# # adding labels
# SG = rep('', times=N) # group label
# SG[1:group1]='1' # 
# SG[(group1+1):N]='2'
# 
# a_per_group=data.frame(SG,S,infect_month)
# c_per_group = a_per_group %>%
#   filter(infect_month > 0 ) %>%
#   group_by(SG,infect_month) %>% 
#   summarise(mean_per_Month=mean(S),N_inf=n())
# # plotting
# barplotU = data.frame(G = c_per_group$SG ,Month = c_per_group$infect_month,N_inf = c_per_group$N_inf)
# 
# SVG = rep('', times=N) #group label
# SVG[1:group3]="3"
# SVG[(group3+1):NV]="4"
# 
# ##############################Bar Plot for Infected##################################
# av_per_group=data.frame(SVG,SV,infect_monthV)
# cv_per_group = av_per_group %>%
#   filter(infect_monthV > 0 ) %>%
#   group_by(SVG,infect_monthV) %>% 
#   summarise(mean_per_Month=mean(SV),N_inf=n())
# barplotV = data.frame(G = cv_per_group$SVG ,Month = cv_per_group$infect_monthV,N_inf= cv_per_group$N_inf)
# 
# # all four groups combined for the bar plot
# barplotALL=rbind(barplotU,barplotV)
# #ggplot(barplotALL, aes(x = Month, y = N_inf, fill = G, colour = G)) + geom_bar(stat = "identity")
# #ggsave("barplotALL.eps", width = 6, height = 4.5, device="eps") 
# write.csv(barplotALL, file = paste0("barplotALL1_0_3_0_95_uncor_",m,".csv"), row.names = FALSE)
# 
# ##############################Bar Plot for Severe##################################
# #################progressed to severe in unvaccinated###################
# as_per_group=data.frame(SG,S,ser_month)
# cs_per_group = as_per_group %>%
#   filter(ser_month > 0 ) %>%
#   group_by(SG,ser_month) %>% 
#   summarise(mean_per_Month=mean(S),N_ser=n())
# barplotSU = data.frame(G = cs_per_group$SG ,Month = cs_per_group$ser_month,N_ser= cs_per_group$N_ser)
# #####progressed to severe in vaccinated###################
# avs_per_group=data.frame(SVG,SV,ser_monthV)
# cvs_per_group = avs_per_group %>%
#   filter(ser_monthV > 0 ) %>%
#   group_by(SVG,ser_monthV) %>% 
#   summarise(mean_per_Month=mean(SV),N_ser=n())
# barplotSV = data.frame(G = cvs_per_group$SVG ,Month = cvs_per_group$ser_monthV,N_ser= cvs_per_group$N_ser)
# 
# barplotALLS=rbind(barplotSU,barplotSV)
# write.csv(barplotALLS, file = paste0("barplotALLS1_0_3_0_95_uncor_",m,".csv"), row.names = FALSE)
# ###############################################################################################
# ##################################calculating VEp###############################
# VEp=1-(dV$NserV/cV$NV)[2:7]/(d$Nser/c$N)[2:7] 
#    return(VEp)
#    })
# }
# 
# set.seed(123) #setting seed keeps ALL 100 simulations the same if you repeat it
# simul_number=100
# VEpbasic <- matrix(NA,nrow=simul_number,ncol=6)
# for (m in 1:simul_number) {
# VEpbasic[m,] <- VEmodelwithbarplots(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,730,0,m)
# }
# write.csv(VEpbasic, file = "VEpbasic_waning730f007_1_0_3_0_95_uncor.csv", row.names = FALSE)
# dropbox_path <- "~/Dropbox/"
# write.csv(VEpbasic, 
#           file = paste0(dropbox_path, "VEpbasic_waning730f007_1_0_3_0_95_uncor.csv"), 
#           row.names = FALSE)
```

```{r figure7EFplots, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}

###########################  Figure 7 Panels E and F plots   #######################################
#Finding the closest to the average curve and it's number

########Plotting all 100 simulation and the one that is closest to average####################

Plotting100<-function(filename){ 
  with(as.list(c(filename)),{
    
# Read the CSV file
file = paste0(filename, ".csv")
data <- read.csv(file)
  
strings <- data
# Calculate the overall average values 
overall_averaged_values <- apply(strings, 2, mean) 

# Function to find the string with the closest values to the overall averaged values 
find_closest_string <- function(strings, overall_averaged_values) { 
  # Calculate the distances between each string and the overall averaged values 
  distances <- apply(strings, 1, function(x) sqrt(sum((x - overall_averaged_values)^2))) 
  # Find the index of the string with the minimum distance 
  closest_string_index <- which.min(distances) 
  return(list(string = strings[closest_string_index, ], index = closest_string_index)) 
  } 
# Find the closest string to the overall averaged values 
closest_string <- find_closest_string(strings, overall_averaged_values) 

pdf(file = paste0(filename, ".pdf"), width = 6.7, height = 5.5, family = "Helvetica", pointsize = 12, onefile = TRUE)
# Plot all strings 
matplot(1:6, t(strings), type = "l", col = "grey", lty = 1, xlab = "Study Month", ylab = "VEp", ylim=c(0,0.8)) 
# Plot the chosen closest string as a thick line 
lines(1:6, closest_string$string, type = "l", lwd = 2) 
#legend("topleft", legend = paste("Closest to Average (VEp#", closest_string$index,")", sep = " "), col = "black", lwd = 2)
dev.off()
     return(closest_string$index)
   })
 }

################finding the closest to averaged VE values for each condition ############ 
allfilenames <- c("VEpbasic_waning730f007uncor0_05_0_7", "VEpbasic_waning730f007_1_0_3_0_95_uncor")

for (k in 1:length(allfilenames)) {
index=Plotting100(allfilenames[k])
file = paste0(allfilenames[k], ".csv")
data=read.csv(file)

# Extract the specific row and convert it to a numeric vector
# Create the variable with the name in the name vector
#the result is a varibale with the name allfilenames[i] that has the simulation values closest to the 100 simulation average
assign(allfilenames[k], as.numeric(data[index,]), envir = .GlobalEnv)
}
print("Figure 7E")
knitr::include_graphics("VEpbasic_waning730f007uncor0_05_0_7.pdf")
print("Figure 7F")
knitr::include_graphics("VEpbasic_waning730f007_1_0_3_0_95_uncor.pdf")
```

\newpage

**Supplemental Figures**

**Supplemental Figure 1.** 

```{r supplementalfigure1ABC, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
###########################  Figure 4 Panel A   ############################################
#Finding the closest to the average curve and it's number

########Plotting all 100 simulation and the one that is closest to average####################

Plotting100<-function(filename){ 
  with(as.list(c(filename)),{
    
# Read the CSV file
file = paste0(filename, ".csv")
data <- read.csv(file)
  
strings <- data
# Calculate the overall average values 
overall_averaged_values <- apply(strings, 2, mean) 

# Function to find the string with the closest values to the overall averaged values 
find_closest_string <- function(strings, overall_averaged_values) { 
  # Calculate the distances between each string and the overall averaged values 
  distances <- apply(strings, 1, function(x) sqrt(sum((x - overall_averaged_values)^2))) 
  # Find the index of the string with the minimum distance 
  closest_string_index <- which.min(distances) 
  return(list(string = strings[closest_string_index, ], index = closest_string_index)) 
  } 
# Find the closest string to the overall averaged values 
closest_string <- find_closest_string(strings, overall_averaged_values) 

pdf(file = paste0(filename, ".pdf"), width = 6.7, height = 5.5, family = "Helvetica", pointsize = 12, onefile = TRUE)
# Plot all strings 
matplot(1:6, t(strings), type = "l", col = "grey", lty = 1, xlab = "Study Month", ylab = "VEp") 
# Plot the chosen closest string as a thick line 
lines(1:6, closest_string$string, type = "l", lwd = 2) 
#legend("topleft", legend = paste("Closest to Average (VEp#", closest_string$index,")", sep = " "), col = "black", lwd = 2)
dev.off()
     return(closest_string$index)
   })
 }

################finding the closest to averaged VE values for each condition ############ 
allfilenames <- c("VEpbasic_waning730f007_0_5x","VEpbasic_waning730f007_1","VEpbasic_waning730f007_2x")

for (k in 1:length(allfilenames)) {
index=Plotting100(allfilenames[k])
file = paste0(allfilenames[k], ".csv")
data=read.csv(file)

# Extract the specific row and convert it to a numeric vector
# Create the variable with the name in the name vector
#the result is a varibale with the name allfilenames[i] that has the simulation values closest to the 100 simulation average
assign(allfilenames[k], as.numeric(data[index,]), envir = .GlobalEnv)
}

print("Supplemental Figure 1A")
knitr::include_graphics("VEpbasic_waning730f007_0_5x.pdf")
print("Supplemental Figure 1B")
knitr::include_graphics("VEpbasic_waning730f007_1.pdf")
print("Supplemental Figure 1C")
knitr::include_graphics("VEpbasic_waning730f007_2x.pdf")
```

The code will use the data from .csv files "susceptible91_2xless.csv", "susceptible1_69.csv", and "susceptible59_2xmore.csv" to plot Supplemental Figure 1 Panels D, E, and F. 

```{r supplementalfigure1DEF, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
# Load necessary libraries
rm(list = ls())
library(readr)   # Reading CSV files
library(tidyr)   # Data tidying functions
library(dplyr)   # Data manipulation functions
library(ggplot2) # Plotting with ggplot2

# Load the CSV file
data <- read.csv("susceptible91_2xless.csv") 
#data <- read.csv("susceptible1_69.csv") #baseline parameters
#data <- read.csv("susceptible59_2xmore.csv")  

# Extract relevant columns (assuming 2nd and 6th columns contain infection dates)
infection_dates <- data[[2]]
infection_dates_vaccinated <- data[[6]]

# Define groups
group1 <- infection_dates[1:14001]
group2 <- infection_dates[14002:200000]
group3 <- infection_dates_vaccinated[1:14001]
group4 <- infection_dates_vaccinated[14002:200000]
total <- rbind(group1, group2, group3, group4)

# Function to calculate the fraction of susceptible people per day
fraction_susceptible <- function(group, days = 1:180) {
  total <- length(group)  # Total individuals in the group

  # Calculate the fraction of susceptible individuals per day
  sapply(days, function(day) {
    still_susceptible <- sum(group == 0 | group > day)  # Count individuals not infected by this day
    still_susceptible / total  # Fraction of susceptible individuals
  })
}

# Apply the function to each group
frac1 <- fraction_susceptible(group1)
frac2 <- fraction_susceptible(group2)
frac3 <- fraction_susceptible(group3)
frac4 <- fraction_susceptible(group4)

# Prepare data for plotting (include day 0 as 1 for all groups)
plot_data <- data.frame(
  Days = rep(0:180, 4),  # Days for x-axis positioning
  Fraction = c(1, frac1, 1, frac2, 1, frac3, 1, frac4),  # Fractions for each group
  Group = factor(rep(1:4, each = 181))  # Group labels
)

# Define color-blind-friendly colors for the groups
ColorBlindFriendly <- c("#E69F00", "#56B4E9", "#800080", "#009E73")
    
# Create the plot using ggplot2
ggplot(plot_data, aes(x = Days, y = Fraction, color = Group)) +
  geom_line(size = 1.5) +  # Make lines thicker with size parameter
  ylim(0, 1) +  # Set y-axis limits between 0 and 1scale_colour_manual(values = ColorBlindFriendly) +
  scale_colour_manual(values = ColorBlindFriendly) +
  scale_x_continuous(
    breaks = seq(0, 180, by = 30),  # Set breaks at every 30 days
    labels = c("0", "1", "2", "3", "4", "5", "6")  # Correct labels
  ) +
  labs(
    #title = "Fraction of Susceptible Individuals in Different Groups",
    x = "Study Month",  # X-axis label
    y = "Fraction of Susceptible",  # Y-axis label
    color = "Group"  # Legend title
  ) +
  theme(
    axis.title.x = element_text(size = 18),  # X-axis label size
    axis.title.y = element_text(size = 18),  # Y-axis label size
    axis.text.x = element_text(size = 18),   # X-axis tick label size
    axis.text.y = element_text(size = 18),   # Y-axis tick label size
    legend.title = element_text(size = 18),  # Legend title size
    legend.text = element_text(size = 18)    # Legend text size
  )
print("Supplemental Figure 1D")
ggsave("Sfigure1D.eps", width = 6, height = 4.5, device = "eps")
#ggsave("Sfigure1E.eps", width = 6, height = 4.5, device = "eps")
#ggsave("Sfigure1F.eps", width = 6, height = 4.5, device = "eps")

```

\newpage

**Supplemental Figure 2.** 

Supplemental Figure 2 Panels A and D are the same as Figure 6A and Figure 7B. 
Supplemental Figure 2 Panels B and E are the same as Figure 6D and Figure 7E. 
The code below is for Supplemental Figure 2 Panel F as an example on how correlation was introduced.
Files "VEpbasic_waning730f007_0_5_0_95_cor.csv" to plot Panel C and "VEpbasic_waning730f007cor0_05_0_7.csv" to plot Panel F are provided.

```{r supplementalFigure2PanelF, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}

rm(list=ls())
library(readr)   # Read tabular data.
library(tidyr)   # Data frame tidying functions.
library(dplyr)   # General data frame manipulation.
library(ggplot2) # Flexible plotting.
# 
# VEmodelwithbarplots<-function(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m){ #m is the number of simulations
#   with(as.list(c(frac1,frac2,alpha1,alpha2,alpha3,alpha4,beta1,beta2,beta3,beta4,VEswaning,VEpwaning,m)),{ # m is the number of the simulation
# # population
# N =200000 #number of unvaccinated people      
# NV=200000 #number of vaccinated people 
# 
# adjust_alpha=1
# adjust_beta=1 #1/0.8 where 0.8 is a mean value for vulnerable group
# 
# #Unvaccinated
# infection_force = 1/180.*adjust_alpha   #4./180. # 30% attack rate over 180 days; to be adjusted
# infection_to_serious_force = 1/10*adjust_beta # base chance to get severe disease if infected
# #Vaccinated
# infection_forceV = 1/180.*adjust_alpha #4./180
# infection_to_serious_forceV = 1/10.*adjust_beta # base chance to get severe disease if infected while being vaccinated
# 
# group1=ceiling(N*frac1)-1 #correction
# group3=ceiling(NV*frac2)-1 #correction
# 
# sample_vulnerableV <- NULL
# sample_immunocompetentV <- NULL
# 
# ###########means 0.7 (vulnerable) and 0.05 (immunocompetent) and a variance of 0.01#############
# alpha_vulnerable <- 14
# beta_vulnerable <- 6
# 
# alpha_immunocompetent <- 0.1875
# beta_immunocompetent <- 3.5625
# 
# # Sample coefficients for both groups for parameter 'beta' in the model
# sample_vulnerableV[1:group3] <- rbeta(group3, alpha_vulnerable, beta_vulnerable)
# sample_immunocompetentV[1:(NV-group3)] <- rbeta((NV-group3), alpha_immunocompetent, beta_immunocompetent)
# 
# #if uncorrelated this is for alpha
# sample_vulnerableV_dif <- NULL
# sample_immunocompetentV_dif <- NULL
# 
# # #case when uncorrelated
# # sample_vulnerableV_dif[1:group3] <- rbeta(group3, alpha_vulnerable, beta_vulnerable)
# # sample_immunocompetentV_dif[1:(NV-group3)] <- rbeta((NV-group3), alpha_immunocompetent, beta_immunocompetent)
# 
# #case when correlated
# sample_vulnerableV_dif[1:group3]=sample_vulnerableV[1:group3]
# sample_immunocompetentV_dif[1:(NV-group3)]=sample_immunocompetentV[1:(NV-group3)]
# 
# alpha3v=mean(sample_vulnerableV_dif)
# alpha4v=mean(sample_immunocompetentV_dif)
# print(alpha3v)
# print(alpha4v)
# beta3v=mean(sample_vulnerableV)
# beta4v=mean(sample_immunocompetentV)
# print(beta3v)
# print(beta4v)
# 
# if (VEswaning>0) {
# VEs_waining= (alpha2-alpha4v)/VEswaning #(alpha2-alpha4)/180.
# VEs_waining_V= (alpha1-alpha3v)/VEswaning #for vulnerable group
# }
# if (VEpwaning>0) {
# VEp_waining= (beta2-beta4v)/VEpwaning 
# VEp_waining_V= (beta1-beta3v)/VEpwaning 
# }
# 
# ###########################block for unvaccinated##################################
# #initializing groups S and SS with different protection
# S = 0.0*(1:N)
# SS = 0.0*(1:N)
# infect=0*(1:N)
# ser=0*(1:N) #severe cases
# 
# #two groups: 'vulnerable'(group1) and 'immunocompetent'
# S[1:group1]=alpha1 # protection against infection in unvaccinated: 1 means base, <1 means more protected
# S[(group1+1):N]=alpha2
# SS[1:group1]=beta1 # protection against severe disease: 1 means base, <1 means more protected
# SS[(group1+1):N]=beta2 
# 
# for (i in 1:180) #i corresponds to the day, j corresponds to the individal/person
# { # cycle per days
#   for (j in 1:N) # cycle per people
#   {
#     if ((infect[j]==0) && (infection_force*S[j] > runif(1))) # infection probability
#     {
#       infect[j]=i #day of infection
#       if (infection_to_serious_force*SS[j] > runif(1)) # progression to severe probability
#       {
#         ser[j]=i #day of becoming severely ill
#       }
#     }
#   }
# }
# # sum_infect=sum(infect>0)
# # sum_ser= sum(ser>0)
# # ratio_inf_to_ser = sum_infect/sum_ser
# 
# #calculate at what month each person was infected; 0 - person is not infected
# infect_month=(infect+30)%/%31 
# 
# a=data.frame(S,infect,infect_month,ser)
# b = a %>% group_by(infect) %>% summarise(mean=mean(S))
# 
# #infect_month=0 in 'c' shows the line for uninfected; if absent -all infected; if '6' is absent - noone infected during month 6
# c = a %>% group_by(infect_month) %>% summarise(mean_per_Month=mean(S),N=n())
# 
# ser_month=(ser+30)%/%30.1
# a_ss=data.frame(S,ser,ser_month)
# ##ser_month=0 in 'd' shows the line for those not progressed to severe disease 
# d = a_ss %>% group_by(ser_month) %>% summarise(SS_per_Month=mean(S),Nser=n())
# 
# ###############################################vaccinated##############################
# #initializing groups S and SS with different protection
# SV = 0.0*(1:NV)
# SSV = 0.0*(1:NV)
# infectV=0*(1:NV)
# serV=0*(1:NV)
# 
# #two groups: 'vulnerable'(group3) and 'immunocompetent'(group 4) in vaccinated
# 
# if (VEswaning==0) {
#   SV[1:group3]=sample_vulnerableV_dif[1:group3] 
#   SV[(group3+1):NV]=sample_immunocompetentV_dif[1:(NV-group3)]
# }
# 
# if (VEpwaning==0) {
# # #two lines below are needed only for stable progression to severe disease#
# # #also need to comment the line SSV=SV below
#   SSV[1:group3]=sample_vulnerableV[1:group3]
#   SSV[(group3+1):NV]=sample_immunocompetentV[1:(NV-group3)]  #beta4
# }
# 
# for (i in 1:180)  { # cycle per days
#   if (VEswaning>0) {
#   SV[1:group3]=sample_vulnerableV_dif[1:group3] #alpha3
#   SV[(group3+1):NV]=sample_immunocompetentV_dif[1:(NV-group3)] #alpha4
#   if(mean(sample_immunocompetentV_dif[1:(NV-group3)])<alpha2){
#   sample_immunocompetentV_dif[1:(NV-group3)]= sample_immunocompetentV_dif[1:(NV-group3)]+VEs_waining
#  for (schet in 1:(NV-group3)){
#     if (sample_immunocompetentV_dif[schet]>1) {sample_immunocompetentV_dif[schet]==1}
#   }
#   }
#   if(mean(sample_vulnerableV_dif[1:group3])<alpha1){
#   sample_vulnerableV_dif[1:group3]= sample_vulnerableV_dif[1:group3]+VEs_waining_V
#     for (schet in 1:group3){
#     if (sample_vulnerableV_dif[schet]>1) {sample_vulnerableV_dif[schet]==1}
#   }
#   }
#   }
#   if (VEpwaning>0) {
#   SSV[(group3+1):NV]=sample_immunocompetentV[1:(NV-group3)]  #beta4
#   SSV[1:group3]=sample_vulnerableV[1:group3]  #beta3
#   if(mean(sample_immunocompetentV[1:(NV-group3)])<beta2){
#   sample_immunocompetentV[1:(NV-group3)]= sample_immunocompetentV[1:(NV-group3)]+VEp_waining
#   for (schet in 1:(NV-group3)){
#     if (sample_immunocompetentV[schet]>1) {sample_immunocompetentV[schet]==1}
#   }
#   }
#   if(mean(sample_vulnerableV[1:group3])<beta1){
#   sample_vulnerableV[1:group3]= sample_vulnerableV[1:group3]+VEp_waining_V
#   for (schet in 1:group3){
#     if (sample_vulnerableV[schet]>1) {sample_vulnerableV[schet]==1}
#   }
#   }
#   }
# 
#   for (j in 1:N) # cycle per people
#   {
#     if ((infectV[j]==0) && (infection_forceV*SV[j] > runif(1))) # infection probability
#     {
#       infectV[j]=i
#       if (infection_to_serious_forceV*SSV[j] > runif(1)) # serious probability
#       {
#         serV[j]=i
#       }
#     }
#   }
# }
# sum_infectV=sum(infectV>0)
# sum_serV= sum(serV>0)
# ratio_inf_to_serV = sum(infectV>0)/sum(serV>0)
# 
# infect_monthV=(infectV+30)%/%31
# aV=data.frame(SV,infectV,infect_monthV,serV)
# bV = aV %>% group_by(infectV) %>% summarise(mean=mean(SV))
# cV = aV %>% group_by(infect_monthV) %>% summarise(mean_per_MonthV=mean(SV),NV=n())
#  
# ser_monthV=(serV+30)%/%30.1
# a_ssV=data.frame(SV,serV,ser_monthV)
# dV = a_ssV %>% group_by(ser_monthV) %>% summarise(SS_per_MonthV=mean(SV),NserV=n())

# ##################################calculating VEp###############################
# VEp=1-(dV$NserV/cV$NV)[2:7]/(d$Nser/c$N)[2:7] 
#    return(VEp)
#    })
# }
# 
# set.seed(123) #setting seed keeps ALL 100 simulations the same if you repeat it
# simul_number=100
# VEpbasic <- matrix(NA,nrow=simul_number,ncol=6)
# for (m in 1:simul_number) {
# VEpbasic[m,] <- VEmodelwithbarplots(0.07,0.07,1,1,1,0.03,1,0.5,1,0.03,730,0,m)
# }
# write.csv(VEpbasic, file = "VEpbasic_waning730f007cor0_05_0_7.csv", row.names = FALSE)

##########################################################################################
########Plotting all 100 simulation and the one that is closest to average####################

Plotting100<-function(filename){ 
  with(as.list(c(filename)),{
    
# Read the CSV file
file = paste0(filename, ".csv")
data <- read.csv(file)
  
strings <- data
# Calculate the overall average values 
overall_averaged_values <- apply(strings, 2, mean) 

# Function to find the string with the closest values to the overall averaged values 
find_closest_string <- function(strings, overall_averaged_values) { 
  # Calculate the distances between each string and the overall averaged values 
  distances <- apply(strings, 1, function(x) sqrt(sum((x - overall_averaged_values)^2))) 
  # Find the index of the string with the minimum distance 
  closest_string_index <- which.min(distances) 
  return(list(string = strings[closest_string_index, ], index = closest_string_index)) 
  } 
# Find the closest string to the overall averaged values 
closest_string <- find_closest_string(strings, overall_averaged_values) 

pdf(file = paste0(filename, ".pdf"), width = 6.7, height = 5.5, family = "Helvetica", pointsize = 12, onefile = TRUE)
# Plot all strings 
matplot(1:6, t(strings), type = "l", col = "grey", lty = 1, xlab = "Study Month", ylab = "VEp") 
# Plot the chosen closest string as a thick line 
lines(1:6, closest_string$string, type = "l", lwd = 2) 
#legend("topleft", legend = paste("Closest to Average (VEp#", closest_string$index,")", sep = " "), col = "black", lwd = 2)
dev.off()
     return(closest_string$index)
   })
 }
################finding the closest to averaged VE values for each condition ############ 
allfilenames <- c("VEpbasic_waning730f007_0_05_0_95_cor", "VEpbasic_waning730f007cor0_05_0_7")

for (k in 1:length(allfilenames)) {
index=Plotting100(allfilenames[k])
file = paste0(allfilenames[k], ".csv")
data=read.csv(file)

# Extract the specific row and convert it to a numeric vector
# Create the variable with the name in the name vector
#the result is a varibale with the name allfilenames[i] that has the simulation values closest to the 100 simulation average
assign(allfilenames[k], as.numeric(data[index,]), envir = .GlobalEnv)
}

print("Supplemental Figure 2C")
knitr::include_graphics("VEpbasic_waning730f007_0_05_0_95_cor.pdf")

print("Supplemental Figure 2F")
knitr::include_graphics("VEpbasic_waning730f007cor0_05_0_7.pdf")

```


